### 2016-12-02 report/paper.tex
Created structure for report

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{balance}
\usepackage[nooneline,hang,it,IT]{subfigure}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{\LARGE TN1008 \\ Advanced Simulation and Visualization of Fluids in Computer Graphics \\ \Large Divergence-Free Smoothed Particle Hydrodynamics}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Ronja Grosz, rongr946\\ Isabell Jansson, isaja187\\ Jonathan Bosson, jonbo665}
\date{\today}
%\authorfooter{
%% insert punctuation at end of each item
%\item
 %Ronja Grosz is a student at Link\"oping University, Sweden, e-mail: rongr946@student.liu.se.
%}

%% Abstract section.
\abstract{ 
}
%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Divergence-free, SPH, divergence correction, density correction.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\firstsection{Introduction}
\maketitle 
%explain the context of the work:
	%What exactly is the problem?
	%What have you created?

\section{Background and Related Work}
%What have people done before?
	%When addressing this problem, when addressing similar problems
%What have you done that makes you approach different?
Based on~\cite{bender}

\section{Method}
%What have you done?
%How did you do it?
%Why did you do it that way?
%May want to mention why you *didn’t* do it in some 
%other possible way
%Clear enough that you could hand the report to another student and they could reproduce the work!
%Probably no need to mention programming language etc
\subsection{Neighbourhood search}
\subsection{Divergence solver}
\subsection{Density solver}
\subsection{Kernel}
\subsection{Navier-stokes}
% Write about non pressure forces, velocity osv..?
\subsection{Adapted time step}
\subsection{Density and alpha factors}

\section{Implementation}
%Details of how it was done
%What hardware
%What software
%Why did you do it that way?
%What limitations, if any, did this present?
%*No code*!!!!!

    % Talk about what we used for plugins, computers etc maybe or should this be in method?

\section{Results}
%What is the end product:
%What does it look like?
%How does one control and interact with it?
%How quick is it?
%What limitations does it have?
%Be self-critical and honest about it - or I will be

\section{Conclusions and Future Work}
%Based on the results and evaluation
%Say what you’ve done right... ...and what you’ve done wrong! - be honest!
%Suggest some possible ways that the work could have been done better
%Suggest some ways that it could be extended and improved by adding more effort

% Use ghost-SPH or similar

\bibliographystyle{abbrv}
\bibliography{./refs}

\end{document}
\grid

### 2016-12-02 particle/Particle.cpp
Fixed it?

@@ -51,6 +51,9 @@ void Particle::setDensity(double newDensity) {
 void Particle::setDensityAdv(double newDensity) {
 	densityAdv = newDensity;
 }
+void Particle::setDerivedDensity(double newDensity) {
+	derivedDensity = newDensity;
+}
 void Particle::setAlpha(double newAlpha) {
 	alpha = newAlpha;
 }
@@ -62,6 +65,7 @@ ivec4 Particle::getCellIndex(){return cellIndex;}
 vector<int>* Particle::getNeighbours(){return neighbours;}
 double Particle::getDensity(){return density;}
 double Particle::getDensityAdv(){return densityAdv;}
+double Particle::getDerivedDensity(){return derivedDensity;}
 double Particle::getAlpha(){return alpha;}
 
 // Update position with current velocity

### 2016-12-02 particle/Particle.h

@@ -51,6 +51,7 @@ class Particle {
 		// Properties
 		double 	density;
 		double  densityAdv;
+		double	derivedDensity;
 		double  alpha; // for kernelfunction (ai)
 
 	public:
@@ -64,6 +65,7 @@ class Particle {
         void setCellIndex(ivec4 cell);
 		void setDensity(double);
 		void setDensityAdv(double);
+		void setDerivedDensity(double);
 		void setAlpha(double);
 
 		// Getters
@@ -73,6 +75,7 @@ class Particle {
         vector<int>* getNeighbours();
 		double getDensity();
 		double getDensityAdv();
+		double getDerivedDensity();
 		double getAlpha();
 		
 		void updatePosition(double elapsedTime);	// Apply the velocity to the position

### 2016-12-02 particle/SPH.cpp

@@ -37,15 +37,15 @@ SPH::SPH()
 	srand(time(0));
 
 	water = new vector<Particle*>(particleCount);
-    cellList = new CellList(dvec3(-0.5, -0.5, -0.5), dvec3(0.5, 0.5, 0.5), H);
+    cellList = new CellList(dvec3(-1.0, -1.0, -1.0), dvec3(1.0, 1.0, 1.0), H);
 
     // Initiate particles
 	for(int i = 0; i < particleCount; ++i) {
 
 		// Random position
-		randX = ((float)rand()/(float)RAND_MAX) * 1.0 - 0.5;
-		randY = ((float)rand()/(float)RAND_MAX) * 1.0 - 0.5;
-		randZ = ((float)rand()/(float)RAND_MAX) * 1.0 - 0.5;
+		randX = ((float)rand()/(float)RAND_MAX) * 2.0 - 1.0;
+		randY = ((float)rand()/(float)RAND_MAX) * 2.0 - 1.0;
+		randZ = ((float)rand()/(float)RAND_MAX) * 2.0 - 1.0;
 
 		// Random velocity
 		randI = (((double)rand()/(double)RAND_MAX) * 0.2) - 0.1;
@@ -54,21 +54,21 @@ SPH::SPH()
 		
 		water->at(i) = new Particle();
 		water->at(i)->setPosition(randX, randY, randZ);
-		water->at(i)->setVelocity(dvec3(randI, randJ, randK));
+		water->at(i)->setVelocity(dvec3(0.0, 0.0, 0.0));
 	}
 
 	for(int i = 0; i < particleCount; ++i) {
 		// Divide particles into cells
         cellList->addParticle(water->at(i), i);
 	}
-
+	
 	// Find neighbours
     for (int i = 0; i < particleCount; ++i) {
         water->at(i)->updateNeighbours(cellList->findNeighbours(water, i));
     }
 
 	// Compute densities and alpha factors
-	calculateDensityAndAlpha();		
+	calculateDensityAndAlpha();	
 
 	createVAO();
 }
@@ -104,9 +104,9 @@ void SPH::loadJson(string fileName) {
 
     // Load particle properties
     particleRadius = params.get<picojson::object>()["particleRadius"].get<double>();
-    //particleMass = params.get<picojson::object>()["particleMass"].get<double>();
+    particleMass = params.get<picojson::object>()["particleMass"].get<double>();
     restDensity = params.get<picojson::object>()["density"].get<double>();
[…]

### 2016-12-02 particle/SPH.h

@@ -35,6 +35,8 @@ class SPH
 		int				particleCount;	
 		double			dT;
 		double 			restDensity;
+		double			avgDensityAdv;
+		double 			avgDerivedDensity;
 
 		vector <Particle*> 	*water;	//this is my vector full of particles
         CellList* cellList;
@@ -60,7 +62,7 @@ class SPH
 		void predictVelocities();
 		dvec3 dirichletBoundary(dvec3, dvec3, dvec3);
 		bool isSolid(dvec4);
-		void calculateDensityAdv(int);
+		void calculateDensityChange();
 		void correctDensityError();
 		void calculateDensityAndAlpha();	
 		void correctDivergenceError();

### 2016-12-02 particle/SPH.cpp
fix on particles getting stuck on edges

@@ -182,11 +182,11 @@ void SPH::predictVelocities() {
 		dPos = vel * dT;
 
 		// Dirichlet Boundary Condition
-		if(isSolid(dvec4(pos.x+dPos.x, pos.y, pos.z, 1.0))) // X
+		if(isSolid(dvec4(pos.x+dPos.x, pos.y, pos.z, 1.0), 0)) // X
 			vel.x = -0.1*vel.x; //0.0;
-		if(isSolid(dvec4(pos.x, pos.y+dPos.y, pos.z, 1.0))) // Y
+		if(isSolid(dvec4(pos.x, pos.y+dPos.y, pos.z, 1.0), 1)) // Y
 			vel.y = -0.1*vel.y; //0.0;
-		if(isSolid(dvec4(pos.x, pos.y, pos.z+dPos.z, 1.0))) // Z
+		if(isSolid(dvec4(pos.x, pos.y, pos.z+dPos.z, 1.0), 2)) // Z
 			vel.z = -0.1*vel.z; //0.0;
 		
 		water->at(i)->setVelocity(vel);
@@ -202,7 +202,7 @@ dvec3 SPH::dirichletBoundary(dvec3 pos, dvec3 dPos, dvec3 vel) {
 }
 
 // Check if (x,y,z) is inside an implicit geometry
-bool SPH::isSolid(dvec4 p) {
+bool SPH::isSolid(dvec4 p, int i) {
 	dmat4 Q = dmat4(0.0);
 
 	// Cylinder
@@ -245,6 +245,22 @@ bool SPH::isSolid(dvec4 p) {
 		Q[3][3] = -geometry.w; // + or -
 	}
 
+
+	switch (i) {
+		case 0: 
+			return (abs(p.x) > 1.0);
+			break;
+		case 1:
+			return (p.y < -1.0);
+			break;
+		case 2:
+			return (abs(p.z) > 1.0);
+			break;
+		default:
+			return false;
+			break;
+	}
+	/*
 	if (max(abs(p.x),abs(p.y),abs(p.z)) > 1.0)
 		return true;
 	// Cube
@@ -255,7 +271,8 @@ bool SPH::isSolid(dvec4 p) {
 			return true;
 	}
 	else
-		return false; //return (dot(p,(Q * p)) < 0.1);
+		return false; //(dot(p,(Q * p)) < 0.1);
+	*/
 }
 
 // Calculate density function

### 2016-12-02 particle/SPH.h

@@ -61,7 +61,7 @@ class SPH
 		void adaptTimestep();
 		void predictVelocities();
 		dvec3 dirichletBoundary(dvec3, dvec3, dvec3);
-		bool isSolid(dvec4);
+		bool isSolid(dvec4, int);
 		void calculateDensityChange();
 		void correctDensityError();
 		void calculateDensityAndAlpha();	

### 2016-12-02 render/Viewport.cpp

@@ -34,7 +34,7 @@ std::string ZeroPadNumber(int num) {
 
 Viewport::Viewport() {
 	phi = 0.0f;
-	theta = M_PI / 4.0f;
+	theta = M_PI / 8.0f;
 	rad = 4.0f;
 	zoomFactor = M_PI;
 	recordTime = deltaTime = currTime = 0.0f;	

### 2016-12-02 particle/SPH.cpp
Small optimisation, divergenceerror only runs 1 time now. need to look into etaV

@@ -164,12 +164,6 @@ void SPH::adaptTimestep() {
 	// make sure dT is less than the maximum timestep
 	if (maxTimestep < dT)
 		dT = maxTimestep;
-/*
-	if (dT > 0.005)
-		dT = 0.005;
-	else if (dT < 0.0005)
-		dT = 0.0005;
-*/
 }
 
 // Predicts the velocity of the particle with its non-pressure forces and dirichlet boundary condition
@@ -285,32 +279,21 @@ void SPH::calculateDensityAndAlpha() {
                 = water->at(i)->getNeighbours()->begin();
                 it != water->at(i)->getNeighbours()->end(); ++it) {
 
+        	// Density
 			water->at(i)->setDensity(water->at(i)->getDensity() + particleMass
                 * water->at(i)->kernel(water->at(*it)->getPosition(), H));
 
-			//cout << "density: " << water->at(i)->getDensity() << endl;
-			// Only need to calc within neighborhood, kernel gradient will be zero otherwise 	
-			/*sum1 += particleMass
-                * water->at(i)->gradientKernel(water->at(*it)->getPosition(), H);
-			
-			sum2 += dot(abs(particleMass
-                * water->at(i)->gradientKernel(water->at(*it)->getPosition(), H)),
-                 abs(particleMass
-                * water->at(i)->gradientKernel(water->at(*it)->getPosition(), H)));*/
 
+			// Alpha sum denominator
 			dvec3 gradPj = particleMass * water->at(i)->gradientKernel(water->at(*it)->getPosition(), H);
 			sumGradPk += dot(gradPj,gradPj); 
 			gradPi += gradPj;
 
 		}
 		sumGradPk += dot(gradPi, gradPi);
-		//cout << i << ": " << water->at(i)->getDensity() << "                                 \n";
-		alpha = std::max(sumGradPk, 1.0e-6); //water->at(i)->getDensity()/(dot(abs(sum1),abs(sum1)) + sum2);
-		if (alpha <= 1.0e-6)
-			cout << alpha << "\n";
+		alpha = std::max(sumGradPk, 1.0e-6); // doesnt really do anything
 		alpha = 1.0/alpha;
 		water->at(i)->setAlpha(alpha); 
-		//cout << water->at(i)->getDensity() << ", " << water->at(i)->getDensityAdv() << "\n";
 
 	}
 }
[…]

### 2016-12-04 render/Viewport.cpp
added support for controlling the gravity direction

@@ -43,7 +43,8 @@ Viewport::Viewport() {
 	frameCount = 0;
 	timeSinceAction = glfwGetTime();
 
-	timeSinceStart = new timer();
+	phiW = 0.0;
+	thetaW = 0.0;
 }
 
 Viewport::~Viewport() {}
@@ -106,6 +107,7 @@ void Viewport::interaction(GLFWwindow *window) {
 	deltaTime = (glfwGetTime() - currTime) / 10;
 	currTime = glfwGetTime();
 
+	// Zoom
 	if (glfwGetKey(window, GLFW_KEY_RIGHT_CONTROL) || glfwGetKey(window, GLFW_KEY_LEFT_CONTROL)) {
 		if (glfwGetKey(window, GLFW_KEY_UP) || glfwGetKey(window, GLFW_KEY_W)) {
 			if (rad > 0.0f)
@@ -115,6 +117,7 @@ void Viewport::interaction(GLFWwindow *window) {
 			rad += deltaTime*zoomFactor;
 		}
 	}
+	// Up/Down
 	else {
 		if (glfwGetKey(window, GLFW_KEY_UP) || glfwGetKey(window, GLFW_KEY_W)) {
 			theta += deltaTime*M_PI / 2.0; // Rotate 90 degrees per second
@@ -126,6 +129,7 @@ void Viewport::interaction(GLFWwindow *window) {
 		}
 	}
 
+	// Left/Right
 	if (glfwGetKey(window, GLFW_KEY_RIGHT) || glfwGetKey(window, GLFW_KEY_D)) {
 		phi -= deltaTime*M_PI / 2.0; // Rotate 90 degrees per second (pi/2)
 		phi = fmod(phi, M_PI*2.0); // Wrap around at 360 degrees (2*pi)
@@ -135,15 +139,37 @@ void Viewport::interaction(GLFWwindow *window) {
 		phi += deltaTime*M_PI / 2.0; // Rotate 90 degrees per second (pi/2)
 		phi = fmod(phi, M_PI*2.0);
 	}
+
+	// Record
 	if (glfwGetKey(window, GLFW_KEY_R) && recordTime > 0.5) {
         record = !record;
         if (record)
             std::cout << "Starting to record..                                 " << std::endl;
         else
             std::cout << "Recorded " << recordTime << " seconds (" 
-        << frameCount/recordTime << "fps) Approximately " << ((double)width*height*3/10000000)*frameCount << " MB\n";
[…]

### 2016-12-04 render/Viewport.h

@@ -13,7 +13,6 @@ Viewport is used as a OpenGL controller.  Viewport is responsible for managing a
 #include <iostream>
 #include <vector>
 #include <math.h>
-#include "boost/timer.hpp"
 
 #ifdef __linux__
 #include "../glm/glm/glm.hpp"
@@ -32,7 +31,6 @@ Viewport is used as a OpenGL controller.  Viewport is responsible for managing a
 class SPH;
 
 using namespace std;
-using namespace boost;
 
 struct rect {
 	float width;
@@ -58,6 +56,9 @@ class Viewport {
 		float theta;
 		float rad;
 
+		float phiW;
+		float thetaW;
+
 		float zoomFactor;
 
 		double deltaTime;
@@ -68,7 +69,6 @@ class Viewport {
 		int frameCount;
 
 		SPH 		*hydro;
-		timer		*timeSinceStart;
 		//CAVIGenerator mov;
 		//BYTE* bm
 

### 2016-12-04 results/merge.sh
Recompiled .sh to 25 frames (this is user based depending on chosen fps in scene_parameters)

@@ -3,7 +3,7 @@
 # -i = image directory, %04d makes 0001 to 9999
 # -crf = quality, 15-25 is good
 
-ffmpeg -r 30 -f image2 -s 640x480 -i frames/frame%04d.png -vcodec libx264 -crf 25  -pix_fmt yuv420p out.mp4;
+ffmpeg -r 25 -f image2 -s 640x480 -i frames/frame%04d.png -vcodec libx264 -crf 25  -pix_fmt yuv420p out.mp4;
 
 cd frames/;
 rm *.png;

### 2016-12-04 particle/SPH.cpp
Added support for gravity control, fixed even playup speed in rendered video, divergence solver working decently, fixed boundary condition with both dirichlet and custom geometry (geometry is based on its gradient in that position)

@@ -19,7 +19,7 @@ SPH is responsible for orginization of a group of smooth particles.
 #include "../json/picojson.h"
 #include "../particle/SPH.h"
 
-#define EPSILON 0.000000000000001f
+#define EPSILON 0.001f
 
 using namespace glm;
 
@@ -28,30 +28,26 @@ SPH::SPH()
 	// load json parameters
 	loadJson("json/scene_parameters.json");
 
-	//scene.readOBJ(sceneName.c_str());
-	//scene.printInfo();
-
 	GLfloat randX, randY, randZ;
-	double randI, randJ, randK; // Velocity vector values
 	
 	srand(time(0));
 
 	water = new vector<Particle*>(particleCount);
-    cellList = new CellList(dvec3(-1.0, -1.0, -1.0), dvec3(1.0, 1.0, 1.0), H);
+    cellList = new CellList(dvec3(-1.0, 0.0, -1.0), dvec3(1.0, 3.0, 1.0), H);
+    kappa = new vector<double>(particleCount);
 
     // Initiate particles
 	for(int i = 0; i < particleCount; ++i) {
 
 		// Random position
 		randX = ((float)rand()/(float)RAND_MAX) * 2.0 - 1.0;
-		randY = ((float)rand()/(float)RAND_MAX) * 2.0 - 1.0;
+		randY = ((float)rand()/(float)RAND_MAX) * 2.0 + 0.0;
 		randZ = ((float)rand()/(float)RAND_MAX) * 2.0 - 1.0;
 
-		// Random velocity
-		randI = (((double)rand()/(double)RAND_MAX) * 0.2) - 0.1;
-		randJ = (((double)rand()/(double)RAND_MAX) * 0.2) - 0.1;
-		randK = (((double)rand()/(double)RAND_MAX) * 0.2) - 0.1;
-		
+		if (alongBoundary(dvec4(randX, randY, randZ, 1.0)) != dvec3(0,0,0)) {
+			randY += 0.5;
+		}
+
 		water->at(i) = new Particle();
 		water->at(i)->setPosition(randX, randY, randZ);
 		water->at(i)->setVelocity(dvec3(0.0, 0.0, 0.0));
@@ -118,14 +114,16 @@ void SPH::loadJson(string fileName) {
 
 // Update positions with a small timestep
 void SPH::simulate() {
-	// Adapt timestep according to CFL condition
-	adaptTimestep();
+	
 
 	// Predict velocities
[…]

### 2016-12-04 particle/SPH.h

@@ -39,6 +39,7 @@ class SPH
 		double 			avgDerivedDensity;
 
 		vector <Particle*> 	*water;	//this is my vector full of particles
+        vector <double> *kappa;
         CellList* cellList;
 		GLuint vao, vbo[2];		// handles to vao and abo
 		
@@ -51,6 +52,9 @@ class SPH
 		double  constantAcceleration;
 		double  maxError;
 		double  maxErrorV;
+		double  maxVelocity;
+
+		dvec3 	gravity;
 
 		std::string sceneName;
 		vec4 geometry;
@@ -60,8 +64,8 @@ class SPH
 
 		void adaptTimestep();
 		void predictVelocities();
-		dvec3 dirichletBoundary(dvec3, dvec3, dvec3);
-		bool isSolid(dvec4, int);
+		void boundaryCondition();
+		dvec3 alongBoundary(dvec4);
 		void calculateDensityChange();
 		void correctDensityError();
 		void calculateDensityAndAlpha();	
@@ -74,7 +78,7 @@ class SPH
 		SPH();
 		~SPH();
 
-		void display();
+		void display(float, float);
 
 };
 

### 2016-12-04 particle/SPH.cpp
Cleanup and added epsilon (frameError) to json file

@@ -19,8 +19,6 @@ SPH is responsible for orginization of a group of smooth particles.
 #include "../json/picojson.h"
 #include "../particle/SPH.h"
 
-#define EPSILON 0.001f
-
 using namespace glm;
 
 SPH::SPH()
@@ -107,9 +105,10 @@ void SPH::loadJson(string fileName) {
     maxError = params.get<picojson::object>()["maxError"].get<double>();
     maxErrorV = params.get<picojson::object>()["maxErrorV"].get<double>();
     
+    frameError = params.get<picojson::object>()["frameError"].get<double>();
     
     cout << "Particle radius: " << particleRadius << "m\n";
-    cout << "Particle mass calculated from radius: " << particleMass*1000.0 << "g\n";
+    cout << "Particle mass: " << particleMass*1000.0 << "g\n";
 }
 
 // Update positions with a small timestep
@@ -121,9 +120,6 @@ void SPH::simulate() {
 
 	// correctDensityError
 	correctDensityError();
-
-	// dirichlet and geometry boundaries
-	boundaryCondition();
 
 	// Update particles position and cell
 	for (int i = 0; i < particleCount; ++i) {
@@ -167,39 +163,36 @@ void SPH::adaptTimestep() {
 
 // Predicts the velocity of the particle with its non-pressure forces and dirichlet boundary condition
 void SPH::predictVelocities() {
-	
 	gravity *= dT;
 	for (int i = 0; i < particleCount; ++i) {
 		water->at(i)->setVelocity(water->at(i)->getVelocity() + gravity);
 	}
 	gravity /= dT;
-	
 }
 
-void SPH::boundaryCondition() {
-	for (int i = 0; i < particleCount; ++i) {
-		dvec3 vel = water->at(i)->getVelocity();
-		dvec3 pos = water->at(i)->getPosition();
-		dvec3 dPos = vel * dT;
-		dvec3 gradP = alongBoundary(dvec4(pos+dPos, 1.0));
+// Dirichlet and geometry boundary conditions. Making sure we don't move particles into solids
+void SPH::boundaryCondition(int i) {
+	dvec3 vel = water->at(i)->getVelocity();
[…]

### 2016-12-04 particle/SPH.h

@@ -52,6 +52,7 @@ class SPH
 		double  constantAcceleration;
 		double  maxError;
 		double  maxErrorV;
+		double  frameError;
 		double  maxVelocity;
 
 		dvec3 	gravity;
@@ -64,7 +65,7 @@ class SPH
 
 		void adaptTimestep();
 		void predictVelocities();
-		void boundaryCondition();
+		void boundaryCondition(int);
 		dvec3 alongBoundary(dvec4);
 		void calculateDensityChange();
 		void correctDensityError();

### 2016-12-05 particle/SPH.cpp
Kappa changes through the iterations, so it shouldnt be outside the while loops

@@ -32,8 +32,7 @@ SPH::SPH()
 
 	water = new vector<Particle*>(particleCount);
     cellList = new CellList(dvec3(-1.0, 0.0, -1.0), dvec3(1.0, 3.0, 1.0), H);
-    kappa = new vector<double>(particleCount);
-
+    
     // Initiate particles
 	for(int i = 0; i < particleCount; ++i) {
 
@@ -312,21 +311,22 @@ void SPH::correctDensityError() {
 	
 	// Lower calculations of ki/kj by adding scaling directly
 	for (int i = 0; i < particleCount; ++i) {
-		double k = std::max((water->at(i)->getDensityAdv() - restDensity) 
-			* water->at(i)->getAlpha() * invdT2, 0.5); 
-		kappa->at(i) = k;
+		water->at(i)->setAlpha(water->at(i)->getAlpha() * invdT2);
 	}
 
 	const double eta = maxError * 0.01 * restDensity;  // maxError is given in percent
 	while ((abs(avgDensityAdv - restDensity) > eta || (iter < 2)) && (iter < 100)) {
 		for (int i = 0; i < particleCount; ++i) {
 			dvec3 sum = dvec3(0.0, 0.0, 0.0);
+			double ki = std::max((water->at(i)->getDensityAdv() - restDensity) 
+			* water->at(i)->getAlpha(), 0.5); 
 
 			for (vector<int>::iterator it = water->at(i)->getNeighbours()->begin();
                 it != water->at(i)->getNeighbours()->end(); ++it) {
-				
-				sum += particleMass 
-					* (kappa->at(i) + kappa->at(*it))
+
+				double kj = std::max((water->at(i)->getDensityAdv() - restDensity) 
+				* water->at(i)->getAlpha(), 0.5); 
+				sum += particleMass * (ki + kj)
 					* water->at(i)->gradientKernel(water->at(*it)->getPosition(), H); 
 			}
 
@@ -336,6 +336,10 @@ void SPH::correctDensityError() {
 		}
 		calculateDensityChange();
 		iter++;
+	}
+	// Change alpha back
+	for (int i = 0; i < particleCount; ++i) {
+		water->at(i)->setAlpha(water->at(i)->getAlpha() * dT*dT);
 	}
 }
 
@@ -349,29 +353,31 @@ void SPH::correctDivergenceError() {
 
 	// Lower calculations of ki/kj by adding scaling directly
 	for (int i = 0; i < particleCount; ++i) {
[…]

### 2016-12-05 particle/SPH.h

@@ -39,7 +39,6 @@ class SPH
 		double 			avgDerivedDensity;
 
 		vector <Particle*> 	*water;	//this is my vector full of particles
-        vector <double> *kappa;
         CellList* cellList;
 		GLuint vao, vbo[2];		// handles to vao and abo
 		

### 2016-12-05 particle/SPH.cpp
A boundingbox is rendered. Vao is created in viewport instead of in sph.

@@ -70,7 +70,8 @@ SPH::SPH()
 	// Compute densities and alpha factors
 	calculateDensityAndAlpha();	
 
-	createVAO();
+	// Generate (one) new Vertex Buffer Object and get the associated id
+	glGenBuffers(2, vbo);
 }
 
 
@@ -491,7 +492,7 @@ void SPH::correctDivergenceError() {
 	}*/
 }
 
-void SPH::display()	{	
+void SPH::display(GLuint vao)	{	
 	GLfloat vertices[particleCount][3];
 	GLfloat colors[particleCount][3];
 
@@ -547,7 +548,6 @@ void SPH::display()	{
     glEnableVertexAttribArray(1);
 
 	glClearColor(0.0f, 0.0f, 0.0f, 1.0f); // Make background black
-	glClear(GL_COLOR_BUFFER_BIT | GL_DEPTH_BUFFER_BIT);
 
 	glBindVertexArray(vao);
 
@@ -558,18 +558,4 @@ void SPH::display()	{
 	glDrawArrays(GL_POINTS, 0, particleCount);
 
 	glDisable(GL_PROGRAM_POINT_SIZE);
-}
-
-void SPH::createVAO () {
-	// A VAO (Vertex Array Object) stores information of a complete rendered object.
-	// It contains all VBOs (Vertex Buffer Objects)
-	// A VBO stores information about the vertices. 
-	// Now we're using two VBOs, one for coordinates and one for colors
-
-	// Allocate and bind Vertex Array Object to the handle vao
-	glGenVertexArrays(1, &vao);
-	glBindVertexArray(vao);
-
-	// Generate (one) new Vertex Buffer Object and get the associated id
-	glGenBuffers(2, vbo);
 }

### 2016-12-05 particle/SPH.h

@@ -40,7 +40,7 @@ class SPH
 
 		vector <Particle*> 	*water;	//this is my vector full of particles
         CellList* cellList;
-		GLuint vao, vbo[2];		// handles to vao and abo
+		GLuint vbo[2];		// handles to vao and abo
 		
 		// particle properties
 		double 	particleRadius;
@@ -56,7 +56,6 @@ class SPH
 		vec4 geometry;
 
 		void loadJson(std::string);
-		void createVAO();
 
 		void adaptTimestep();
 		void predictVelocities();
@@ -66,16 +65,13 @@ class SPH
 		void correctDensityError();
 		void calculateDensityAndAlpha();	
 		void correctDivergenceError();
-		void simulate();	//gets neighboring particels and calls their getForceAtPoint, applyForce...
-		
-		
+		void simulate();	//gets neighboring particels and calls their getForceAtPoint, applyForce...	
 		
 	public:
 		SPH();
 		~SPH();
 
-		void display();
-
+		void display(GLuint vao);
 };
 
 #endif

### 2016-12-05 render/BoundingBox.h

@@ -1,0 +1,21 @@
+/*************************************************************************
+Name:	    Isabell Jansson, Jonathan Bosson, Ronja Grosz
+File name: 	BoundingBox.h
+
+*************************************************************************/
+
+#include "GL/glew.h"
+#include "GLFW/glfw3.h"
+
+class BoundingBox {
+	private:
+		int scale;
+		GLuint vbo_bb[2];
+
+	public:
+		BoundingBox();
+		~BoundingBox();
+
+		void draw(GLuint vao, double scale);
+		
+};

### 2016-12-05 render/Viewport.cpp

@@ -87,6 +87,9 @@ void Viewport::init(void) {
 }
 
 void Viewport::initWorld() {
+	glGenVertexArrays(1, &vao); //maybe move to init()
+	glBindVertexArray(vao);
+
 	Viewport::hydro = new SPH();	//this is the object that will manage all of the particles
 }
 
@@ -238,9 +241,9 @@ int Viewport::start(int argc, char** argv) {
         glUniform3fv(locationCa, 1, glm::value_ptr(cam));
         */
 
-		hydro->display();
-		
-
+        hydro->display(vao);
+        boundingBox.draw(vao, 1.2);
+        
 		// Save the frame
 		if (record) {
 			frameCount++;

### 2016-12-05 render/Viewport.h

@@ -14,6 +14,8 @@ Viewport is used as a OpenGL controller.  Viewport is responsible for managing a
 #include <vector>
 #include <math.h>
 #include "boost/timer.hpp"
+#include "../render/BoundingBox.h"
+
 
 #ifdef __linux__
 #include "../glm/glm/glm.hpp"
@@ -47,6 +49,8 @@ class Viewport {
 		~Viewport();
 		virtual int start(int argc, char** argv);			//This functions is called to start the program.
 
+		GLuint vao;
+
 	private:
 		// void detectCollisions(stellarBody*,stellarBody*);	//I chose to implement a collision detection algorithm
 		int width;
@@ -68,6 +72,7 @@ class Viewport {
 		int frameCount;
 
 		SPH 		*hydro;
+		BoundingBox	boundingBox;
 		timer		*timeSinceStart;
 		//CAVIGenerator mov;
 		//BYTE* bm

### 2016-12-05 report/paper.tex
started introduction and writing on cell list

@@
 	%What have you created?
+    Smoothed particle hydrodynamics, \textit{SPH}, is a method that was first implemented in 1977 for astrophysical simulations by Gingold et al.~\cite{firstSPH}.
+    Since then SPH has become a popular method for complex water simulations.
+    SPH is a mesh-free Lagrangian method where the particles move in space and change physical properties as time progresses.
+    % More about Lagrangian and maybe navier stokes?
+
+    % We used this article
+    In this paper we are going to introduce the results from reproducing the divergence-free smoothed particle hydrodynamics method introduced by Bender et al.~\cite{bender}.
+    It is a method which corrects the divergence error, aiming for a divergence-free velocity field which is needed for an incompressible fluid.
+    For the solution to be divergence-free the density has to be constant over time.
+
+    % Write about other solutions? Why is this solution the BEST!!?
 
@@
     %What have you done that makes you approach different?
-Based on~\cite{bender}
+
+    % Bridson
+    ~\cite{bridson}.
+
+    % Write about Benders article
 
@@
 \subsection{Neighbourhood search}
+    Since SPH only considers a finite amount of neighbouring particles, it is important to keep track of every particles neighbours.
+    Searching through all particles for neighbours within the cutoff distance $H$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
+    The cutoff distance $H$ is the kernel smoothing radius.
+    To fasten this up a cell list was implemented.
+    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $H$.
+    When finding the neighbour of particle $i$, only the neighbouring cells have to be searched for particles within the cutoff distance, see figure~\ref{fig:cellList}.
+
+    \begin{figure}
+    \includegraphics[width=\linewidth]{img/CellLists2.png}
+    \caption{Finding the neighbours for the filled in particle $i$ by looking through all neighbouring cells, including its own cell, for particles within the cutoff distance $H$}
+    \label{fig:cellList}
+    \end{figure}
+
+    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates for the cells and the fourth dimension is for storing the particles belonging to the cell.
+    The amount of cells are decided by dividing the scene into cells of length $H$.
+    The particles are then assigned to a cell.
+    If the particle moves out of its cell it is then assigned to the new cell.
+
+    \begin{equation} \label{eq:assignCell}
+
+    \end{equation}
+
 \subsection{Divergence solver}
@@
 \subsection{Density and alpha factors}
+\subsection{Screen space fluid rendering}
 
@@
 
-\bibliographystyle{abbrv}
+%\bibliographystyle{abbrv}
+\bibliographystyle{ieeetr}
 \bibliography{./refs}

### 2016-12-05 particle/SPH.cpp
remove some unnecessary code

@@ -281,7 +281,6 @@ void SPH::calculateDensityAndAlpha() {
 		alpha = std::max(sumGradPk, 1.0e-6); // doesnt really do anything?
 		alpha = 1.0/alpha;
 		water->at(i)->setAlpha(alpha); 
-
 	}
 }
 
@@ -428,29 +427,11 @@ void SPH::display(float phiW, float thetaW, GLuint vao)	{
 		vertices[i][0] = water->at(i)->getPosition().x;
 		vertices[i][1] = water->at(i)->getPosition().y;
 		vertices[i][2] = water->at(i)->getPosition().z;
-		
 
 		colors[i][0] = 0.7 - water->at(i)->getPosition().x;
 		colors[i][1] = 0.7 - water->at(i)->getPosition().y;
 		colors[i][2] = 0.7 - water->at(i)->getPosition().z;
-		
-		/*
-		colors[i][0] = 0.9;
-		colors[i][1] = 0.9;
-		colors[i][2] = 0.9;
-		
-		dvec3 vel = water->at(i)->getVelocity();
-		double mag = dot(vel, vel) / maxVelocity;
-
-		if (mag > 0.5) {
-			colors[i][0] = mag;
-			colors[i][1] = 0.1;
-			colors[i][2] = 0.1;
-		}
-		*/
-
 	}
-
 
 	// Bind the first VBO as being the active buffer and storing vertex attributes (coordinates)
     glBindBuffer(GL_ARRAY_BUFFER, vbo[0]);

### 2016-12-05 render/Viewport.cpp

@@ -85,11 +85,12 @@ void Viewport::init(void) {
 	glCullFace(GL_BACK);
 	glFrontFace(GL_CW);
 	glPolygonMode(GL_FRONT_AND_BACK, GL_POINT); // with?
+
+	glGenVertexArrays(1, &vao); //maybe move to init()
+	glBindVertexArray(vao);
 }
 
 void Viewport::initWorld() {
-	glGenVertexArrays(1, &vao); //maybe move to init()
-	glBindVertexArray(vao);
 
 	Viewport::hydro = new SPH();	//this is the object that will manage all of the particles
 }
@@ -280,8 +281,7 @@ int Viewport::start(int argc, char** argv) {
 			// Free resources
 			FreeImage_Unload(image);
 			delete [] pixels;
-		}
-		
+		}	
 		
 		glfwSwapBuffers(window);			//swap the buffer
     }

### 2016-12-05 particle/SPH.cpp
shaderwork to make ssfr

@@ -245,7 +245,7 @@ dvec3 SPH::alongBoundary(dvec4 p) {
 	}*/
 	if (dot(p,(Q * p)) < 0.0) {
 		dmat3x4 Qsub = Q;
-		gradP = 2 * Q * p;
+		gradP = 2 * Qsub * p;
 	}
 	return dvec3(gradP);
 }
@@ -474,7 +474,7 @@ void SPH::display(float phiW, float thetaW)	{
 
 	glEnable(GL_PROGRAM_POINT_SIZE); // Enable gl_PointSize in vertex shader
 
-	glPolygonMode(GL_FRONT_AND_BACK, GL_POINT);
+	glPolygonMode(GL_BACK, GL_POINT);
 
 	glDrawArrays(GL_POINTS, 0, particleCount);
 

### 2016-12-05 render/Viewport.cpp

@@ -232,7 +232,7 @@ int Viewport::start(int argc, char** argv) {
     locationMV = glGetUniformLocation(phongShader.programID, "MV");
 	locationP = glGetUniformLocation(phongShader.programID, "P");
 	//locationLight = glGetUniformLocation(phongShader.programID, "lightPos");
-	//locationCamera = glGetUniformLocation(phongShader.programID, "camPos");
+	locationCamera = glGetUniformLocation(phongShader.programID, "camPos");
 
     // Let's get started!
     while (!glfwWindowShouldClose(window)) {
@@ -258,11 +258,11 @@ int Viewport::start(int argc, char** argv) {
         //convert viewMatrix to float
         glUniformMatrix4fv(locationMV, 1, GL_FALSE, glm::value_ptr(viewMatrix));
 
-        /*        
+                
         // send in light and camera location to glsl (not done in shaders yet)
-        glUniform3fv(locationL, 1, glm::value_ptr(light));
-        glUniform3fv(locationCa, 1, glm::value_ptr(cam));
-        */
+        //glUniform3fv(locationL, 1, glm::value_ptr(light));
+        glUniform3fv(locationCamera, 1, glm::value_ptr(cameraPosition));
+        
         
 		hydro->display(phiW, thetaW);
 		

### 2016-12-05 particle/SPH.cpp
Qsub doesnt work with plane, why? Changed to Q, which works and gives reasonable results but shouldnt be the gradient

@@ -187,7 +187,7 @@ void SPH::boundaryCondition(int i) {
 
 	// Boundary of geometry
 	if (gradP != dvec3(0.0, 0.0, 0.0))
-		vel = gradP + gravity*dT;
+		vel = gradP;// + gravity*dT;
 
 	water->at(i)->setVelocity(vel);
 }
@@ -253,7 +253,7 @@ dvec3 SPH::alongBoundary(dvec4 p) {
 			Qsub[3][i] = 0;
 			Qsub[3][i] = 0;
 		}
-		gradP = 2.0 * Qsub * p;
+		gradP = 2.0 * Q * p;
 	}
 	return dvec3(gradP);
 }

### 2016-12-05 render/BoundingBox.cpp

@@ -12,7 +12,7 @@ BoundingBox::BoundingBox(){}
 
 BoundingBox::~BoundingBox() {}
 
-void BoundingBox::draw(GLuint vao, double scale) {
+void BoundingBox::draw(GLuint vao, float scale) {
 	GLfloat vertices[72] = {-scale, scale, scale, 	// 1
 							-scale, scale, -scale, // 2
 

### 2016-12-05 report/paper.tex
Equation compiles
going to unstash refs.bib
kernel function is described in the report, added ref to Monaghan

@@
 
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{mathptmx}
@@
     The amount of cells are decided by dividing the scene into cells of length $H$.
-    The particles are then assigned to a cell.
+    The particles are then assigned to a cell according to Equation ~\ref{eq:assignCell}.
     If the particle moves out of its cell it is then assigned to the new cell.
@@
     \begin{equation} \label{eq:assignCell}
+        sjdfkljsf
+    \end{equation}
 
+% Should be mentioned early since it is used in mostly of the algorithms
+\subsection{Kernel}
+    A kernel function is used to simulate how particle-particle interactions decrease with the distance between the current particle and its neighbours.
+In SPH simulations this is an approximation of the Gaussian kernel function.
+Different kernels have been tested in previous works i.e.
+the poly6 kernel, the spiky kernel and the cubic spline kernel.
+According to Bender et.
+al.~\cite{bender} the cubic spline kernel presented by Monaghan~\cite{monaghan} was used.
+The kernel is described by Equation~\ref{eq:kernel}, where $q(x)=\frac{\left \| x \right \|}{h}$, $x$ is the distance between the current particle and a neighbour particle and $h$ is the support radius for the kernel.
+Particles further away than the support radius will not affect the current particle.
+
+    \begin{equation} \label{eq:kernel}
+        W_h(q(x)) =  \frac{1}{\pi h^3} \left\{\begin{matrix}
+        1 - \frac{3}{2}q^2 + \frac{3}{4}q^3 & 0 \leqslant q < 1 \\ 
+        \\
+        \frac{1}{4}(2-q)^3 & 1 \leqslant q < 2\\ 
+        \\
+        0 & q \geqslant 2
+      \end{matrix}\right.
     \end{equation}
+
+    The algorithm does also require the kernel gradient.
+
+    It is important to use the same kernel function for both $W_h$ and $\nabla W_h$ to get the prediction and the correction step to be compatible to each other.
 
@@
 \subsection{Density solver}
-\subsection{Kernel}
+
 \subsection{Navier-stokes}

### 2016-12-05 report/paper.tex
Put in some code

--- 2016-12-05 report/paper.tex
@@
 \usepackage[nooneline,hang,it,IT]{subfigure}
+
+% Our packages
+\usepackage{listings} % To write code in latex
+\usepackage{color}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\footnotesize,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2,
+    literate={\ \ }{{\ }}1
+}
+
+\lstset{style=mystyle}
 
@@
 %Probably no need to mention programming language etc
+
+\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
+function performSimulation
+    for all particles $i$ do //Init neighbourhoods
+        find neighbourhoods $N_i(0)$
+    for all particles $i$ do //Init $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$
+        compute densities $\rho_i(0)$
+        compute factors $\alpha_i(0)$
+    while (t < $t_{max}$) do //Start simulation loop
+        adapt time step $\Delta{t}$ 
+        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
+            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
+        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
+        for all particles $i$ do //Update positions
+            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
+        for all particles $i$ do //Update neighbourhoods
+            find neighbourhoods $N_i(t + \Delta{t})$
+        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
+            compute densities $\rho_i(0)$
+            compute factors $\alpha_i(0)$
+        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
+
+\end{lstlisting}
+
 \subsection{Neighbourhood search}
@@
     The amount of cells are decided by dividing the scene into cells of length $H$.
-    The particles are then assigned to a cell.
+    The particles are then assigned to a cell according to equation~ref{eq:assignCell}.
     If the particle moves out of its cell it is then assigned to the new cell.
@@
     \begin{equation} \label{eq:assignCell}
-
+        sjdfkljsf
     \end{equation}

### 2016-12-07 particle/SPH.cpp
small change to boundary

@@ -249,9 +249,6 @@ dvec3 SPH::alongBoundary(dvec4 p) {
 
 		for (int i = 0; i < 4; ++i) {
 			Qsub[3][i] = 0;
-			Qsub[3][i] = 0;
-			Qsub[3][i] = 0;
-			Qsub[3][i] = 0;
 		}
 		gradP = 2.0 * Q * p;
 	}

### 2016-12-08 particle/SPH.cpp
Rendering spheres now with shading, no ssfr

@@ -27,7 +27,7 @@ SPH::SPH()
 	loadJson("json/scene_parameters.json");
 
 	GLfloat randX, randY, randZ;
-	
+	maxVelocity = constantAcceleration * maxTimestep;	
 	srand(time(0));
 
 	water = new vector<Particle*>(particleCount);
@@ -178,11 +178,11 @@ void SPH::boundaryCondition(int i) {
 	dvec3 gradP = alongBoundary(dvec4(pos+dPos, 1.0));
 
 	// Dirichlet
-	if(pos.x+dPos.x < -1.2 || pos.x+dPos.x > 1.2) // X
+	if(pos.x+dPos.x < -1.0 || pos.x+dPos.x > 1.0) // X
 		vel.x = 0.0;
-	if(pos.y+dPos.y < -1.2) // Y
+	if(pos.y+dPos.y < -1.0) // Y
 		vel.y = 0.0;
-	if(pos.z+dPos.z < -1.2 || pos.z+dPos.z > 1.2) // Z
+	if(pos.z+dPos.z < -1.0 || pos.z+dPos.z > 1.0) // Z
 		vel.z = 0.0;
 
 	// Boundary of geometry
@@ -245,15 +245,17 @@ dvec3 SPH::alongBoundary(dvec4 p) {
 			return true;
 	}*/
 	if (dot(p,(Q * p)) < 0.0) {
-		dmat4 Qsub = Q;// = dmat3x4(dvec3(Q[0]), dvec3(Q[1]), dvec3(Q[2], dvec3(Q[3])));
-
+		dmat4 Qsub = Q;
 		for (int i = 0; i < 4; ++i) {
 			Qsub[3][i] = 0;
 			Qsub[3][i] = 0;
 			Qsub[3][i] = 0;
 			Qsub[3][i] = 0;
 		}
-		gradP = 2.0 * Qsub * p; // doesnt work with plane
+		if (sceneName == "plane")
+			gradP = 2.0 * Q * p;
+		else 
+			gradP = 2.0 * Qsub * p;
 	}
 	return dvec3(gradP);
 }
@@ -391,7 +393,7 @@ void SPH::correctDivergenceError() {
 
 void SPH::display(float phiW, float thetaW, GLuint vao)	{	
 	GLfloat vertices[particleCount][3];
-	GLfloat colors[particleCount][3];
+	GLfloat velocities[particleCount][3];
 
[…]

### 2016-12-08 particle/SPH.h

@@ -31,6 +31,9 @@ using namespace boost;
 
 class SPH
 {
+	public:
+		double  maxVelocity;
+
 	protected:
 		int				particleCount;	
 		double			dT;
@@ -52,7 +55,7 @@ class SPH
 		double  maxError;
 		double  maxErrorV;
 		double  frameError;
-		double  maxVelocity;
+		
 
 		dvec3 	gravity;
 

### 2016-12-08 render/Viewport.cpp

@@ -187,7 +187,7 @@ int Viewport::start(int argc, char** argv) {
 					  0.0f, 0.0f, -0.2f, 0.0f };
 	GLint locationP;
 	GLint locationMV;
-	GLint locationLight;
+	GLint locationVel;
 	GLint locationCamera;
 
 	glm::mat4 viewMatrix;
@@ -235,7 +235,7 @@ int Viewport::start(int argc, char** argv) {
     //link variables to shader
     locationMV = glGetUniformLocation(phongShader.programID, "MV");
 	locationP = glGetUniformLocation(phongShader.programID, "P");
-	//locationLight = glGetUniformLocation(phongShader.programID, "lightPos");
+	locationVel = glGetUniformLocation(phongShader.programID, "maxVelocity");
 	locationCamera = glGetUniformLocation(phongShader.programID, "camPos");
 
     // Let's get started!
@@ -262,10 +262,13 @@ int Viewport::start(int argc, char** argv) {
         //convert viewMatrix to float
         glUniformMatrix4fv(locationMV, 1, GL_FALSE, glm::value_ptr(viewMatrix));
 
-        glUniform3fv(locationCamera, 1, glm::value_ptr(cameraPosition));
-        
+        //glUniform3fv(locationCamera, 1, glm::value_ptr(cameraPosition));
+        float maxVel = (float) hydro->maxVelocity;
+        float* maxVelP = &maxVel;
+        glUniform1fv(locationVel, 1, maxVelP);
+
         hydro->display(phiW, thetaW, vao);
-        boundingBox.draw(vao, 1.2f);
+        boundingBox.draw(vao, 1.0f);
         
 		// Save the frame
 		if (record) {
@@ -277,7 +280,7 @@ int Viewport::start(int argc, char** argv) {
 			glPixelStorei(GL_UNPACK_ALIGNMENT, 1);
 			glReadPixels(0, 0, width, height, GL_RGB, GL_UNSIGNED_BYTE, pixels);
 			// Convert to FreeImage format & save to file
-			FIBITMAP* image = FreeImage_ConvertFromRawBits(pixels, width, height, 3 * width, 24, 0x0000FF, 0xFF0000, 0x00FF00, false);
+			FIBITMAP* image = FreeImage_ConvertFromRawBits(pixels, width, height, 3 * width, 24, 0xFF0000, 0x00FF00, 0x0000FF, false);
 			FreeImage_Save(FIF_PNG, image, fileName.c_str(), 0);
 
 			// Free resources

### 2016-12-08 results/merge.sh

@@ -3,7 +3,7 @@
 # -i = image directory, %04d makes 0001 to 9999
 # -crf = quality, 15-25 is good
 
-ffmpeg -r 60 -f image2 -s 640x480 -i frames/frame%04d.png -vcodec libx264 -crf 25  -pix_fmt yuv420p out.mp4;
+ffmpeg -r 100 -f image2 -s 640x480 -i frames/frame%04d.png -vcodec libx264 -crf 25  -pix_fmt yuv420p out.mp4;
 
 cd frames/;
 rm *.png;

### 2016-12-08 particle/SPH.cpp
boundingDimension in json

@@ -95,7 +95,7 @@ void SPH::loadJson(string fileName) {
 	    params.get<picojson::object>()["geometry"]
     		.get<picojson::object>()["r"].get<double>());
 
-
+    boundaryDimension = params.get<picojson::object>()["boundaryDimension"].get<double>();
     // Load particle properties
     particleRadius = params.get<picojson::object>()["particleRadius"].get<double>();
     particleMass = params.get<picojson::object>()["particleMass"].get<double>();
@@ -166,6 +166,7 @@ void SPH::predictVelocities() {
 	gravity *= dT;
 	for (int i = 0; i < particleCount; ++i) {
 		water->at(i)->setVelocity(water->at(i)->getVelocity() + gravity);
+		boundaryCondition(i);
 	}
 	gravity /= dT;
 }
@@ -178,11 +179,11 @@ void SPH::boundaryCondition(int i) {
 	dvec3 gradP = alongBoundary(dvec4(pos+dPos, 1.0));
 
 	// Dirichlet
-	if(pos.x+dPos.x < -1.0 || pos.x+dPos.x > 1.0) // X
+	if(pos.x+dPos.x < -boundaryDimension || pos.x+dPos.x > boundaryDimension) // X
 		vel.x = 0.0;
-	if(pos.y+dPos.y < -1.0) // Y
+	if(pos.y+dPos.y < -boundaryDimension) // Y
 		vel.y = 0.0;
-	if(pos.z+dPos.z < -1.0 || pos.z+dPos.z > 1.0) // Z
+	if(pos.z+dPos.z < -boundaryDimension || pos.z+dPos.z > boundaryDimension) // Z
 		vel.z = 0.0;
 
 	// Boundary of geometry
@@ -339,9 +340,22 @@ void SPH::correctDensityError() {
 					* water->at(i)->gradientKernel(water->at(*it)->getPosition(), H); 
 			}
 
+			//neumann
+			dvec3 vel = water->at(i)->getVelocity() - (sum * dT);
+			dvec3 pos = water->at(i)->getPosition();
+			dvec3 dPos = vel * dT;
+
+			if(pos.x+dPos.x < -boundaryDimension || pos.x+dPos.x > boundaryDimension) // X
+				sum.x = 0.0;
+			if(pos.y+dPos.y < -boundaryDimension) // Y
+				sum.y = 0.0;
+			if(pos.z+dPos.z < -boundaryDimension || pos.z+dPos.z > boundaryDimension) // Z
+				sum.z = 0.0;
+
[…]

### 2016-12-08 particle/SPH.h

@@ -55,6 +55,7 @@ class SPH
 		double  maxError;
 		double  maxErrorV;
 		double  frameError;
+		double  boundaryDimension;
 		
 
 		dvec3 	gravity;

### 2016-12-08 render/Viewport.cpp

@@ -15,6 +15,7 @@ Viewport is used as a OpenGL controller.  Viewport is responsible for managing a
 
 
 #include "../particle/SPH.h"
+#include "../json/picojson.h"
 #include "../render/Viewport.h"
 #include "Shader.h"
 
@@ -238,6 +239,12 @@ int Viewport::start(int argc, char** argv) {
 	locationVel = glGetUniformLocation(phongShader.programID, "maxVelocity");
 	locationCamera = glGetUniformLocation(phongShader.programID, "camPos");
 
+	picojson::value param;
+	ifstream paramStream ("json/scene_parameters.json");
+	paramStream >> param;
+	float boundaryDimension = (float)(param.get<picojson::object>()["boundaryDimension"].get<double>());
+
+
     // Let's get started!
     while (!glfwWindowShouldClose(window)) {
     	glfwPollEvents();
@@ -249,7 +256,7 @@ int Viewport::start(int argc, char** argv) {
 
 		setupPerspective(window, P);
 		interaction(window);
-		cameraPosition = glm::vec3(0.0f, 0.0f, rad);
+		cameraPosition = glm::vec3(0.0f, -0.5f, rad);
 
 
 		// I is the normal Identity matrix
@@ -268,7 +275,9 @@ int Viewport::start(int argc, char** argv) {
         glUniform1fv(locationVel, 1, maxVelP);
 
         hydro->display(phiW, thetaW, vao);
-        boundingBox.draw(vao, 1.0f);
+
+        
+        boundingBox.draw(vao, boundaryDimension);
         
 		// Save the frame
 		if (record) {

### 2016-12-08 particle/SPH.cpp
cleanup of commented code

@@ -229,7 +229,6 @@ dvec3 SPH::alongBoundary(dvec4 p) {
 		Q[3][2] = -geometry.w/2.0;
 	}
 
-
 	// Hyperboloid
 	else if (sceneName == "hyperboloid") {
 		Q[0][0] = geometry.w;
@@ -238,13 +237,6 @@ dvec3 SPH::alongBoundary(dvec4 p) {
 		Q[3][3] = -geometry.w; // + or -
 	}
 		
-	// Cube
-	/*if (sceneName == "cube") {
-		if (max(abs(p.x),abs(p.y+1.0),abs(p.z)) > geometry.w)
-			return false;
-		else
-			return true;
-	}*/
 	if (dot(p,(Q * p)) < 0.0) {
 		dmat4 Qsub = Q;
 		for (int i = 0; i < 4; ++i) {

### 2016-12-08 particle/Particle.h
using dirichlet on all times after velocity change seems to work decently. Got boundaryParticle code commented in case

@@ -54,7 +54,10 @@ class Particle {
 		double	derivedDensity;
 		double  alpha; // for kernelfunction (ai)
 
+
 	public:
+		int     id;
+		
 		Particle();
 		//Particle(const Particle&);
 		~Particle();

### 2016-12-08 particle/SPH.cpp

@@ -29,16 +29,89 @@ SPH::SPH()
 	GLfloat randX, randY, randZ;
 	maxVelocity = constantAcceleration * maxTimestep;	
 	srand(time(0));
+	double step = 3.0 * particleRadius / boundaryDimension;
 
-	water = new vector<Particle*>(particleCount);
-    cellList = new CellList(dvec3(-1.0, 0.0, -1.0), dvec3(1.0, 3.0, 1.0), H);
+	water = new vector<Particle*>(particleCount); //+(5*pow(ceil(1 + 2*boundaryDimension/step),2)));
+    cellList = new CellList(
+    	dvec3(-2*boundaryDimension, -2*boundaryDimension, -2*boundaryDimension), 
+    	dvec3(2*boundaryDimension, 2*boundaryDimension+3.0, 2*boundaryDimension), H);
     
+    // Create boundary particles
+    idx = 0;
+/*
+    // Bottom
+    randY = -boundaryDimension - particleRadius;
+    for (double i = -boundaryDimension-0.1; i < boundaryDimension+0.1; i+=step) {
+    	for (double j = -boundaryDimension-0.1; j < boundaryDimension+0.1; j+=step) {
+    		water->at(idx) = new Particle();
+			water->at(idx)->setPosition(i, randY, j);
+			water->at(idx)->setVelocity(dvec3(0.0, 0.0, 0.0));
+			water->at(idx)->id = 1;
+			
+			idx++;
+    	}
+    }
+
+    // Front
+    randZ = boundaryDimension + 0.1;
+    for (double i = -boundaryDimension-0.1; i < boundaryDimension+0.1; i+=step) {
+    	for (double j = -boundaryDimension-0.1; j < -boundaryDimension+0.2; j+=step) {
+    		water->at(idx) = new Particle();
+			water->at(idx)->setPosition(i, j, randZ);
+			water->at(idx)->setVelocity(dvec3(0.0, 0.0, 0.0));
+			water->at(idx)->id = 1;
+
+			idx++;
+    	}
+    }
+    // Back
+    randZ = -boundaryDimension - 0.1;
+    for (double i = -boundaryDimension-0.1; i < boundaryDimension+0.1; i+=step) {
+    	for (double j = -boundaryDimension-0.1; j < -boundaryDimension+0.2; j+=step) {
+    		water->at(idx) = new Particle();
+			water->at(idx)->setPosition(i, j, randZ);
[…]

### 2016-12-08 particle/SPH.h

@@ -35,7 +35,8 @@ class SPH
 		double  maxVelocity;
 
 	protected:
-		int				particleCount;	
+		int				particleCount;
+		int 			idx;	
 		double			dT;
 		double 			restDensity;
 		double			avgDensityAdv;
@@ -56,7 +57,6 @@ class SPH
 		double  maxErrorV;
 		double  frameError;
 		double  boundaryDimension;
-		
 
 		dvec3 	gravity;
 

### 2016-12-13 report/paper.tex
gradient kernel in report
changed template for tex

@@
-\documentclass[journal]{vgtc}                % final (journal style)
-%\documentclass[review,journal]{vgtc}         % review (journal style)
-%\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
-%\documentclass[electronic,journal]{vgtc}     % electronic version, journal
+\title{\Large Advanced Simulation and Visualization of Fluids\\ in Computer Graphics \\ \large Divergence-Free Smoothed Particle Hydrodynamics}
 
-%% Uncomment one of the lines above depending on where your paper is
-%% in the conference process. ``review'' and ``widereview'' are for review
-%% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier. Further, ``electronic'' includes
-%% hyperreferences for more convenient online viewing.
+\author{\begin{tabular}{ccc}
+    Isabell Jansson & Ronja Grosz  & Jonathan Bosson \\
+    \small isaja187@student.liu.se &\small rongr946@student.liu.se &\small jonbo665@student.liu.se \\
+\end{tabular}}
 
-%% Please use one of the ``review'' options in combination with the
-%% assigned online id (see below) ONLY if your paper uses a double blind
-%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
-%% in the past.
+\date{\today}
 
-%% Please note that the use of figures other than the optional teaser is not permitted on the first page
-%% of the journal version.  Figures should begin on the second page and be
-%% in CMYK or Grey scale format, otherwise, colour shifting may occur
-%% during the printing process.  Papers submitted with figures other than the optional teaser on the
-%% first page will be refused.
-
-%% These three lines bring in essential packages: ``mathptmx'' for Type 1
-%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
-%% for proper handling of the times font family.
-
-\usepackage{mathptmx}
-\usepackage{graphicx}
-\usepackage{times}
-\usepackage{balance}
-\usepackage[nooneline,hang,it,IT]{subfigure}
+\documentclass[12pt, twocolumn]{article}
 
 % Our packages
+\usepackage{graphicx}   % Images
+\usepackage{amsmath}    % Math
+\usepackage{amssymb}    % Math
 \usepackage{listings}   % To write code in latex
-\usepackage{color}
+\usepackage{color}      % Custom colors
 
@@
 \lstset{style=mystyle}
-
-%% We encourage the use of mathptmx for consistent usage of times font
-%% throughout the proceedings. However, if you encounter conflicts
-%% with other math-related packages, you may want to disable it.
-
-%% This turns references into clickable hyperlinks.
-\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
-\hypersetup{
-  pdfauthor = {},
-  pdftitle = {},
-  pdfsubject = {},
-  pdfkeywords = {},
-  colorlinks=true,
-  linkcolor= black,
-  citecolor= black,
-  pageanchor=true,
-  urlcolor = black,
-  plainpages = false,
-  linktocpage
-}
-
-%% If you are submitting a paper to a conference for review with a double
-%% blind reviewing process, please replace the value ``0'' below with your
-%% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
-
-%% declare the category of your paper, only shown in review mode
-\vgtccategory{Research}
-
-%% allow for this line if you want the electronic option to work properly
-\vgtcinsertpkg
-
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in an IEEE VGTC sponsored conference.}
-
-%% Paper title.
-
-\title{\LARGE TN1008 \\ Advanced Simulation and Visualization of Fluids in Computer Graphics \\ \Large Divergence-Free Smoothed Particle Hydrodynamics}
-
-%% This is how authors are specified in the journal style
-
-%% indicate IEEE Member or Student Member in form indicated below
-\author{Ronja Grosz, rongr946\\ Isabell Jansson, isaja187\\ Jonathan Bosson, jonbo665}
-\date{\today}
-%\authorfooter{
-%% insert punctuation at end of each item
-%\item
- %Ronja Grosz is a student at Link\"oping University, Sweden, e-mail: rongr946@student.liu.se.
-%}
-
-%% Abstract section.
-\abstract{ 
-}
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Divergence-free, SPH, divergence correction, density correction.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% END our packages
 
 \begin{document}
+\twocolumn[
+\begin{@twocolumnfalse}
+\maketitle
 
-\firstsection{Introduction}
-\maketitle 
+\begin{abstract}
+    This is the paper's abstract \ldots
+    \\{\bf Index Terms:} Divergence-free, SPH, divergence correction, density correction.
+\end{abstract}
+
+\end{@twocolumnfalse}
+]
+
+\section{Introduction}
     %explain the context of the work:
@@
         correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
-
 \end{lstlisting}
@@
     The amount of cells are decided by dividing the scene into cells of length $H$.
-    The particles are then assigned to a cell according to equation~ref{eq:assignCell}.
+    The particles are then assigned to a cell according to Equation ~\ref{eq:assignCell}.
     If the particle moves out of its cell it is then assigned to the new cell.
@@
 
+% Should be mentioned early since it is used in mostly of the algorithms
+\subsection{Kernel}
+    A kernel function is used to simulate how particle-particle interactions decrease with the distance between the current particle and its neighbours.
+In SPH simulations this is an approximation of the Gaussian kernel function.
+Different kernels have been tested in previous works i.e.
+the poly6 kernel, the spiky kernel and the cubic spline kernel.
+According to Bender et.
+al.~\cite{bender} the cubic spline kernel presented by Monaghan~\cite{monaghan} was used.
+The kernel is described by Equation~\ref{eq:kernel}, where $q(x)=\frac{\left \| x \right \|}{h}$, $x$ is the distance between the current particle and a neighbour particle and $h$ is the support radius for the kernel.
+Particles further away than the support radius will not affect the current particle.
+
+    \begin{equation} \label{eq:kernel}
+        W_h(q(x)) =  \frac{1}{\pi h^3} \left\{\begin{matrix}
+        1 - \frac{3}{2}q^2 + \frac{3}{4}q^3 & 0 \leqslant q < 1 \\ 
+        \\
+        \frac{1}{4}(2-q)^3 & 1 \leqslant q < 2\\ 
+        \\
+        0 & q \geqslant 2
+      \end{matrix}\right.
+    \end{equation}
+
+    The algorithm does also require the kernel gradient.
+To reduce the computational effort and memory requirements Bender et.
+al.~\cite{bender} introduce a scalar function $g(q) = \frac{\partial W_h}{\partial q} \cdot \frac{1}{h \left \| x \right \|}$.
+The gradient kernel is then calculated by $\partial W_h(q(x)) = x \cdot g(x)$.
+The gradient kernel is described by Equation ~\ref{eq:gradKernel}.
+
+    \begin{equation} \label{eq:gradKernel}
+        \partial W_h(q(x)) =  x \cdot \frac{1}{h \left \| x \right \|} \cdot \frac{1}{\pi h^3} \left\{\begin{matrix}
+        - 3q + \frac{9}{4}q^2 & 0 \leqslant q < 1 \\ 
+        \\
+        - \frac{3}{4}(2-q)^2 & 1 \leqslant q < 2\\ 
+        \\
+        0 & q \geqslant 2
+      \end{matrix}\right.
+    \end{equation}
+    % maybe specify x here as well
+
+    It is important to use the same kernel function for both $W_h$ and $\nabla W_h$ to get the prediction and the correction step to be compatible to each other.
+
 \subsection{Divergence solver}
 \subsection{Density solver}
-\subsection{Kernel}
+
 \subsection{Navier-stokes}
@@
 
-%\bibliographystyle{abbrv}
 \bibliographystyle{ieeetr}
@@
 \end{document}
-\grid

### 2016-12-13 render/Viewport.cpp

@@ -263,8 +263,8 @@ int Viewport::start(int argc, char** argv) {
 		viewMatrix = glm::make_mat4(I);
         // Translate a bit down and backwards
 		viewMatrix = viewMatrix * glm::translate(-cameraPosition)
-        * glm::rotate(theta, glm::vec3(1.0f, 0.0f, 0.0f))
-        * glm::rotate(phi, glm::vec3(0.0f, 1.0f, 0.0f));
+            * glm::rotate(theta, glm::vec3(1.0f, 0.0f, 0.0f))
+            * glm::rotate(phi, glm::vec3(0.0f, 1.0f, 0.0f));
 
         //convert viewMatrix to float
         glUniformMatrix4fv(locationMV, 1, GL_FALSE, glm::value_ptr(viewMatrix));

### 2016-12-14 report/paper.tex
write about cfl

@@
     Smoothed particle hydrodynamics, \textit{SPH}, is a method that was first implemented in 1977 for astrophysical simulations by Gingold et al.~\cite{firstSPH}.
-    Since then SPH has become a popular method for complex water simulations.
-    SPH is a mesh-free Lagrangian method where the particles move in space and change physical properties as time progresses.
+    Since then SPH has become a popular method for complex fluid simulations.
+    SPH is a mesh-free Lagrangian method where the fluid is split into discrete sets defined as particles, which move in space and change physical properties as time progresses.
     % More about Lagrangian and maybe navier stokes?
@@
     % We used this article
-    In this paper we are going to introduce the results from reproducing the divergence-free smoothed particle hydrodynamics method introduced by Bender et al.~\cite{bender}.
+    In this paper we are going to introduce the results from reproducing the divergence-free smoothed particle hydrodynamics method presented by Bender et al.~\cite{bender}.
     It is a method which corrects the divergence error, aiming for a divergence-free velocity field which is needed for an incompressible fluid.
@@
 \subsection{Adapted time step}
+
+% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
+In order to assure a stable simulation is it necessary to adapt the time step depending on the motion.
+The quicker the particles move the smaller time steps are taken such that the numerical error of the explicit time integration is negligible.
+
 \subsection{Density and alpha factors}

### 2016-12-14 report/paper.tex
Adapted timestep

@@
-\title{\Large Advanced Simulation and Visualization of Fluids\\ in Computer Graphics \\ \large Divergence-Free Smoothed Particle Hydrodynamics}
+\title{\vspace{-2cm} \LARGE Divergence-Free Smoothed Particle Hydrodynamics \\ \large Advanced Simulation and Visualization of Fluids in Computer Graphics}
 
 \author{\begin{tabular}{ccc}
-    Isabell Jansson & Ronja Grosz  & Jonathan Bosson \\
-    \small isaja187@student.liu.se &\small rongr946@student.liu.se &\small jonbo665@student.liu.se \\
+    Isabell Jansson\thanks{isaja187@student.liu.se} & Ronja Grosz\thanks{rongr946@student.liu.se}  & Jonathan Bosson\thanks{jonbo665@student.liu.se}
 \end{tabular}}
@@
 
-\documentclass[12pt, twocolumn]{article}
+\documentclass[11pt, twocolumn]{article}
 
@@
 \usepackage{color}      % Custom colors
+\usepackage{abstract}   % Onecolabstract, saythanks
+\usepackage[margin=1.2in]{geometry}
+
+\usepackage{bm}
+\newcommand{\vect}[1]{\bm{#1}} % bold vector
+
+\setlength{\columnsep}{20px} % Length between twocolumn
 
@@
 \twocolumn[
-\begin{@twocolumnfalse}
 \maketitle
 
-\begin{abstract}
-    This is the paper's abstract \ldots
-    \\{\bf Index Terms:} Divergence-free, SPH, divergence correction, density correction.
-\end{abstract}
-
-\end{@twocolumnfalse}
+\begin{onecolabstract}
+    \noindent This is the paper's abstract \ldots
+    \vspace{1em}\\{\bf Index Terms:} Divergence-free, SPH, divergence correction, density correction.\vspace{1em}
+\end{onecolabstract}
 ]
+\saythanks
 
@@
 % CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
-In order to assure a stable simulation is it necessary to adapt the time step depending on the motion.
-The quicker the particles move the smaller time steps are taken such that the numerical error of the explicit time integration is negligible.
+In order to assure a stable simulation it is necessary to adapt the time step depending on the motion. 
+The quicker the particles move the smaller time steps are taken such that the numerical 
+error of the explicit time integration is negligible.
+For this is a condition proposed by Courant–Friedrichs–Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
+The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by the maximum velocity $\vect{V_{max}}$ in the scene.
+
+\begin{equation}\label{eq:cfl}
+\Delta t \leq \frac{0.4 \cdot d}{\vect{V_{max}}}
+\end{equation}
+
+Each rendered frame however have a constant time step in order to ensure no speed change in the playback of the rendered video.
+This is decided from the user side as a desired fps parameter.
+This means that the simulation loop can iterate multiple times within each frame adaptively depending on the maximum velocity.
 

### 2016-12-14 report/paper.tex
changed cell list image, writingt on cellist

--- 2016-12-13 report/paper.tex
@@
 \usepackage{color}      % Custom colors
+\usepackage{abstract}   % Onecolabstract, saythanks
+\usepackage[margin=1.2in]{geometry}
+
+\setlength{\columnsep}{20px} % Length between twocolumn
 
@@
 \begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
-function performSimulation
+function performFluidSimulation
     for all particles $i$ do //Init neighbourhoods
@@
         for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
-            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
+            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_g/m_i$
         correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
@@
 \subsection{Neighbourhood search}
-    Since SPH only considers a finite amount of neighbouring particles, it is important to keep track of every particles neighbours.
+    Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles neighbours.
     Searching through all particles for neighbours within the cutoff distance $H$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
-    The cutoff distance $H$ is the kernel smoothing radius.
-    To fasten this up a cell list was implemented.
+    The cutoff distance $H$ is the kernels smoothing radius.
+    To make this faster a cell list was implemented.
     A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $H$.
+    Each cell spans on a surface in space.
+    A particle belongs to a cell if it is inside the space the cell occupy.
     When finding the neighbour of particle $i$, only the neighbouring cells have to be searched for particles within the cutoff distance, see figure~\ref{fig:cellList}.
 
-    \begin{figure}
-    \includegraphics[width=\linewidth]{img/CellLists2.png}
+    \begin{figure}[h!]\centering
+    \begin{centering}
+    \includegraphics[width=0.7\linewidth]{img/CellList.png}
     \caption{Finding the neighbours for the filled in particle $i$ by looking through all neighbouring cells, including its own cell, for particles within the cutoff distance $H$}
     \label{fig:cellList}
+    \end{centering}
     \end{figure}
 
-    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates for the cells and the fourth dimension is for storing the particles belonging to the cell.
+    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates and the fourth dimension is for storing the particles belonging to the cell.
     The amount of cells are decided by dividing the scene into cells of length $H$.
-    The particles are then assigned to a cell according to Equation ~\ref{eq:assignCell}.
-    If the particle moves out of its cell it is then assigned to the new cell.
+    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $vect{x}$ is the position of the particle and $vect{s}_min$ is the lowest position in the scene space.
+    If the particle moves out of its cell it is reassigned to the new cell it is inside.
 
     \begin{equation} \label{eq:assignCell}
-        sjdfkljsf
+        \frac{\lceil{pos - s_{min}}\rceil}{H}
     \end{equation}

### 2016-12-14 report/paper.tex

@@
 \usepackage[margin=1.2in]{geometry}
+
+\usepackage{bm}
+\newcommand{\vect}[1]{\bm{#1}} % bold vector
 
@@
 \subsection{Adapted time step}
+
+% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
+In order to assure a stable simulation it is necessary to adapt the time step depending on the motion. 
+The quicker the particles move the smaller time steps are taken such that the numerical 
+error of the explicit time integration is negligible.
+For this is a condition proposed by Courant–Friedrichs–Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
+The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by the maximum velocity $\vect{V_{max}}$ in the scene.
+
+\begin{equation}\label{eq:cfl}
+\Delta t \leq \frac{0.4 \cdot d}{\vect{V_{max}}}
+\end{equation}
+
+Each rendered frame however have a constant time step in order to ensure no speed change in the playback of the rendered video.
+This is decided from the user side as a desired fps parameter.
+This means that the simulation loop can iterate multiple times within each frame adaptively depending on the maximum velocity.
+
 \subsection{Density and alpha factors}

### 2016-12-14 report/paper.tex
navier-stokes equations

--- 2016-12-14 report/paper.tex
@@
 \usepackage[margin=1.2in]{geometry}
-
-\usepackage{bm}
-\newcommand{\vect}[1]{\bm{#1}} % bold vector
 
@@
     Smoothed particle hydrodynamics, \textit{SPH}, is a method that was first implemented in 1977 for astrophysical simulations by Gingold et al.~\cite{firstSPH}.
-    Since then SPH has become a popular method for complex fluid simulations.
-    SPH is a mesh-free Lagrangian method where the fluid is split into discrete sets defined as particles, which move in space and change physical properties as time progresses.
+    Since then SPH has become a popular method for complex water simulations.
+    SPH is a mesh-free Lagrangian method where the particles move in space and change physical properties as time progresses.
     % More about Lagrangian and maybe navier stokes?
@@
     % We used this article
-    In this paper we are going to introduce the results from reproducing the divergence-free smoothed particle hydrodynamics method presented by Bender et al.~\cite{bender}.
+    In this paper we are going to introduce the results from reproducing the divergence-free smoothed particle hydrodynamics method introduced by Bender et al.~\cite{bender}.
     It is a method which corrects the divergence error, aiming for a divergence-free velocity field which is needed for an incompressible fluid.
@@
 
-    \begin{equation} \label{eq:gradKernel}
-        \partial W_h(q(x)) =  x \cdot \frac{1}{h \left \| x \right \|} \cdot \frac{1}{\pi h^3} \left\{\begin{matrix}
+    \begin{multline} \label{eq:gradKernel}
+        \partial W_h(q(x)) = \\ x \cdot \frac{1}{h \left \| x \right \|} \cdot \frac{1}{\pi h^3} \left\{\begin{matrix}
         - 3q + \frac{9}{4}q^2 & 0 \leqslant q < 1 \\ 
@@
       \end{matrix}\right.
-    \end{equation}
+    \end{multline}
     % maybe specify x here as well
@@
 
+\subsection{Navier-Stokes}
+    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
+The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
+
+    \begin{equation} \label{eq:nsDensity}
+        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
+    \end{equation}
+
+    \begin{equation} \label{eq:nsVelocity}
+        \frac{D\textbf{v}}{Dt} = \frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
+    \end{equation}
+
+    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
+The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
+Due to the incompressibility, the partial derivate of the density with respect to the time is zero which implies the equivalence of the equations in~\ref{eq:nsDensity} where the second part $\bigtriangledown \cdot \textbf{v} = 0$ induce that the field is free from divergence.
+
 \subsection{Divergence solver}
@@
 
-\subsection{Navier-stokes}
 % Write about non pressure forces, velocity osv..?
 \subsection{Adapted time step}
-
-% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
-In order to assure a stable simulation it is necessary to adapt the time step depending on the motion. 
-The quicker the particles move the smaller time steps are taken such that the numerical 
-error of the explicit time integration is negligible.
-For this is a condition proposed by Courant–Friedrichs–Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
-The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by the maximum velocity $\vect{V_{max}}$ in the scene.
-
-\begin{equation}\label{eq:cfl}
-\Delta t \leq \frac{0.4 \cdot d}{\vect{V_{max}}}
-\end{equation}
-
-Each rendered frame however have a constant time step in order to ensure no speed change in the playback of the rendered video.
-This is decided from the user side as a desired fps parameter.
-This means that the simulation loop can iterate multiple times within each frame adaptively depending on the maximum velocity.
-
 \subsection{Density and alpha factors}

### 2016-12-14 report/paper.tex
finished writing on neighbourhood

--- 2016-12-14 report/paper.tex
@@
     The amount of cells are decided by dividing the scene into cells of length $H$.
-    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $vect{x}$ is the position of the particle and $vect{s}_min$ is the lowest position in the scene space.
+    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_min$ is the lowest position of the scene space.
     If the particle moves out of its cell it is reassigned to the new cell it is inside.
@@
     \begin{equation} \label{eq:assignCell}
-        \frac{\lceil{pos - s_{min}}\rceil}{H}
+        \frac{\lceil{\vect{x} - \vect{s}_{min}}\rceil}{H}
     \end{equation}

### 2016-12-14 report/paper.tex
divergence error start

--- 2016-12-14 report/paper.tex
@@
 
-    \begin{multline} \label{eq:gradKernel}
-        \partial W_h(q(x)) = \\ x \cdot \frac{1}{h \left \| x \right \|} \cdot \frac{1}{\pi h^3} \left\{\begin{matrix}
+    \begin{equation} \label{eq:gradKernel}
+        \partial W_h(q(x)) =  x \cdot \frac{1}{h \left \| x \right \|} \cdot \frac{1}{\pi h^3} \left\{\begin{matrix}
         - 3q + \frac{9}{4}q^2 & 0 \leqslant q < 1 \\ 
@@
       \end{matrix}\right.
-    \end{multline}
+    \end{equation}
     % maybe specify x here as well
@@
 
-\subsection{Navier-Stokes}
-    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
-The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
+\subsection{Divergence-Free solver}
 
-    \begin{equation} \label{eq:nsDensity}
-        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
-    \end{equation}
+\begin{comment}
+\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
+function correctDivergenceError
+    for all particles $i$ do
+        calculateDensityChange $\frac{Dp}{Dt}_i$
+    for all particles $i$ do //Scale $\color{codegreen}a_i$
+        $a_i = \frac{a_i}{\Delta t}$
+    while (t < $t_{max}$) do //Start simulation loop
+        adapt time step $\Delta{t}$ 
+        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
+            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
+        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
+        for all particles $i$ do //Update positions
+            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
+        for all particles $i$ do //Update neighbourhoods
+            find neighbourhoods $N_i(t + \Delta{t})$
+        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
+            compute densities $\rho_i(0)$
+            compute factors $\alpha_i(0)$
+        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
+\end{lstlisting}
+\end{comment}
 
-    \begin{equation} \label{eq:nsVelocity}
-        \frac{D\textbf{v}}{Dt} = \frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
-    \end{equation}
+The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{Dp}{Dt} = 0$, which is equivalent to navier-stokes divergence free condition given in equation %\ref{eq:navier}
 
-    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
-The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
-Due to the incompressibility, the partial derivate of the density with respect to the time is zero which implies the equivalence of the equations in~\ref{eq:nsDensity} where the second part $\bigtriangledown \cdot \textbf{v} = 0$ induce that the field is free from divergence.
-
-\subsection{Divergence solver}
 \subsection{Density solver}
 
+\subsection{Navier-stokes}
 % Write about non pressure forces, velocity osv..?
@@
 The quicker the particles move the smaller time steps are taken such that the numerical 
-error of the explicit time integration is negligible.
-For this is a condition proposed by Courant–Friedrichs–Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
-The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by the maximum velocity $\vect{V_{max}}$ in the scene.
+error of the explicit time integration is negligible. For this is a condition proposed by 
+Courant–Friedrichs–Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}. 
+The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
+the maximum velocity $\vect{V_{max}}$ in the scene.
 

### 2016-12-14 report/paper.tex

--- 2016-12-14 report/paper.tex
@@
 
-\subsection{Divergence solver}
+\subsection{Divergence-Free solver}
+
+\begin{comment}
+\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
+function correctDivergenceError
+    for all particles $i$ do
+        calculateDensityChange $\frac{Dp}{Dt}_i$
+    for all particles $i$ do //Scale $\color{codegreen}a_i$
+        $a_i = \frac{a_i}{\Delta t}$
+    while (t < $t_{max}$) do //Start simulation loop
+        adapt time step $\Delta{t}$ 
+        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
+            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
+        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
+        for all particles $i$ do //Update positions
+            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
+        for all particles $i$ do //Update neighbourhoods
+            find neighbourhoods $N_i(t + \Delta{t})$
+        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
+            compute densities $\rho_i(0)$
+            compute factors $\alpha_i(0)$
+        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
+\end{lstlisting}
+\end{comment}
+
+The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{Dp}{Dt} = 0$, which is equivalent to navier-stokes divergence free condition given in equation %\ref{eq:navier}
+
 \subsection{Density solver}

### 2016-12-14 report/paper.tex

--- 2016-12-14 report/paper.tex
@@
 
-\subsection{Navier-stokes}
 % Write about non pressure forces, velocity osv..?

### 2016-12-14 report/paper.tex
Fixed some equation scaling

@@
     Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles neighbours.
-    Searching through all particles for neighbours within the cutoff distance $H$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
-    The cutoff distance $H$ is the kernels smoothing radius.
+    Searching through all particles for neighbours within the cutoff distance $h$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
+    The cutoff distance $h$ is the kernels smoothing radius.
     To make this faster a cell list was implemented.
-    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $H$.
+    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $h$.
     Each cell spans on a surface in space.
@@
     \includegraphics[width=0.7\linewidth]{img/CellList.png}
-    \caption{Finding the neighbours for the filled in particle $i$ by looking through all neighbouring cells, including its own cell, for particles within the cutoff distance $H$}
+    \caption{Finding the neighbours for the filled in particle $i$ by looking through all neighbouring cells, including its own cell, for particles within the cutoff distance $h$}
     \label{fig:cellList}
@@
     The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates and the fourth dimension is for storing the particles belonging to the cell.
-    The amount of cells are decided by dividing the scene into cells of length $H$.
+    The amount of cells are decided by dividing the scene into cells of length $h$.
     The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_min$ is the lowest position of the scene space.
@@
     \begin{equation} \label{eq:assignCell}
-        \frac{\lceil{\vect{x} - \vect{s}_{min}}\rceil}{H}
+        \frac{\lceil{\vect{x} - \vect{s}_{min}}\rceil}{h}
     \end{equation}
@@
     \begin{equation} \label{eq:kernel}
+        \resizebox{0.88\linewidth}{!}{$
         W_h(q(x)) =  \frac{1}{\pi h^3} \left\{\begin{matrix}
@@
         \end{matrix}\right.
+        $}
     \end{equation}
@@
 
-    \begin{multline} \label{eq:gradKernel}
-        \partial W_h(q(x)) = \\ x \cdot \frac{1}{h \left \| x \right \|} \cdot \frac{1}{\pi h^3} \left\{\begin{matrix}
+    \begin{equation} \label{eq:gradKernel}
+        \resizebox{0.88\linewidth}{!}{$
+        \partial W_h(q(x)) =  x \frac{1}{h \left \| x \right \|} \frac{1}{\pi h^3} \left\{\begin{matrix}
         - 3q + \frac{9}{4}q^2 & 0 \leqslant q < 1 \\ 
@@
         \end{matrix}\right.
-    \end{multline}
+        $}
+    \end{equation}
     % maybe specify x here as well
@@
 
-\subsection{Navier-Stokes}
-    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
-The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
-
-    \begin{equation} \label{eq:nsDensity}
-        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
-    \end{equation}
-
-    \begin{equation} \label{eq:nsVelocity}
-        \frac{D\textbf{v}}{Dt} = \frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
-    \end{equation}
-
-    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
-The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
-Due to the incompressibility, the partial derivate of the density with respect to the time is zero which implies the equivalence of the equations in~\ref{eq:nsDensity} where the second part $\bigtriangledown \cdot \textbf{v} = 0$ induce that the field is free from divergence.
-
 \subsection{Divergence solver}
@@
 
+\subsection{Navier-stokes}
 % Write about non pressure forces, velocity osv..?

### 2016-12-14 report/paper.tex
navier stokes finnished

--- 2016-12-14 report/paper.tex
@@
 The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
-Due to the incompressibility, the partial derivate of the density with respect to the time is zero which implies the equivalence of the equations in~\ref{eq:nsDensity} where the second part $\bigtriangledown \cdot \textbf{v} = 0$ induce that the field is free from divergence.
+The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
+From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
+From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
+In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
+The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
+To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
 

### 2016-12-14 report/paper.tex

--- 2016-12-14 report/paper.tex
@@
 
+\subsection{Navier-Stokes}
+    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
+The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
+
+    \begin{equation} \label{eq:nsDensity}
+        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
+    \end{equation}
+
+    \begin{equation} \label{eq:nsVelocity}
+        \frac{D\textbf{v}}{Dt} = \frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
+    \end{equation}
+
+    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
+The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
+The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
+From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
+From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
+In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
+The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
+To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
+
 \subsection{Divergence solver}

### 2016-12-14 report/paper.tex
density in density and alpha

@@
 % Should be mentioned early since it is used in mostly of the algorithms
-\subsection{Kernel}
+\subsection{Kernel} \label{kernel}
     A kernel function is used to simulate how particle-particle interactions decrease with the distance between the current particle and its neighbours.
@@
 To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
+Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.
 
-\subsection{Divergence solver}
-\subsection{Density solver}
+\subsection{Divergence solver} \label{divergenceSolver}
+\subsection{Density solver} \label{densitySolver}
 
@@
 \subsection{Density and alpha factors}
+    The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}.
+This implies that neighbouring particles closer to the current particle $x_i$ will have a greater affect on the density.
+
+    \begin{equation}\label{eq:density}
+        \rho_i = \sum_j \frac{m_j}{\rho_j} \rho_j W_{ij} = \sum_j m_j W_{ij}
+    \end{equation}
+
 \subsection{Screen space fluid rendering}

### 2016-12-14 report/paper.tex
density solver

@@
     \begin{equation} \label{eq:nsVelocity}
-        \frac{D\textbf{v}}{Dt} = \frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
+        \frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
     \end{equation}
@@
 To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
+Bridson~\cite{bridson} uses splitting on Equation~\ref{eq:nsVelocity} which induce $\frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p$ which will be used further on in the simulation.
 Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.
@@
 \subsection{Density solver} \label{densitySolver}
+    This solver aims to minimize the density error caused by the numerical integration in the divergence solver.
+The density error is determined by comparing the actual density to the rest density, $\rho - \rho_0$.
+There are several density solvers presented in previous works, but the solver used in this project was presented by Bender et.
+al.~\cite{bender} which uses the fact that $\alpha_i$ has been calculated before the solver which reduces the computations in the solver.
+The solver will be active while the average of the predicted densities for every particle differs from the rest density with an error greater than a constant \textit{maxError}.
+For every particle a predicted density $\rho_i^*$ is calculated through an Euler integration step by Equation~\ref{eq:predDensity}.
+
+    \begin{equation} \label{eq:predDensity}
+        \resizebox{0.88\linewidth}{!}{$
+        \begin{matrix}
+        \rho_i^* = \rho_i + \Delta t \frac{D\rho_i}{Dt} \\ 
+        \\
+        \Leftrightarrow \\ 
+        \\
+        \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
+        (\textbf{v}_i^* - \textbf{v}_j^*) \bigtriangledown W_{ij}
+        \end{matrix}
+        $}
+    \end{equation}
+
+    Since the stiffness parameter $\kappa_i$ is derived from the pressure forces, updating the stiffness parameter will correct the density error.
+The stiffness parameter is determined by Equation~\ref{stiffnessDensity}.
+
+    \begin{equation} \label{stiffnessDensity}
+    \kappa_i = \frac{1}{\Delta t^2}(\rho_i^*-\rho_0)\alpha_i
+    \end{equation}
+
+    The velocity is updated according to the the divergence solver. %Det här behöver kanske utvecklas, väntar med det tills stycket om divergence solvern är klar, innehåller samma ekvationer delvis
 

### 2016-12-14 report/paper.tex
divergencefree solver

--- 2016-12-14 report/paper.tex
@@
     Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles neighbours.
-    Searching through all particles for neighbours within the cutoff distance $h$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
-    The cutoff distance $h$ is the kernels smoothing radius.
+    Searching through all particles for neighbours within the cutoff distance $H$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
+    The cutoff distance $H$ is the kernels smoothing radius.
     To make this faster a cell list was implemented.
-    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $h$.
+    A cell list is a data structure that is divided into cells that have a length larger or equals to the cutoff distance $H$.
     Each cell spans on a surface in space.
@@
     \includegraphics[width=0.7\linewidth]{img/CellList.png}
-    \caption{Finding the neighbours for the filled in particle $i$ by looking through all neighbouring cells, including its own cell, for particles within the cutoff distance $h$}
+    \caption{Finding the neighbours for the filled in particle $i$ by looking through all neighbouring cells, including its own cell, for particles within the cutoff distance $H$}
     \label{fig:cellList}
@@
     The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates and the fourth dimension is for storing the particles belonging to the cell.
-    The amount of cells are decided by dividing the scene into cells of length $h$.
+    The amount of cells are decided by dividing the scene into cells of length $H$.
     The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_min$ is the lowest position of the scene space.
@@
     \begin{equation} \label{eq:assignCell}
-        \frac{\lceil{\vect{x} - \vect{s}_{min}}\rceil}{h}
+        \frac{\lceil{\vect{x} - \vect{s}_{min}}\rceil}{H}
     \end{equation}
@@
     \begin{equation} \label{eq:kernel}
-        \resizebox{0.88\linewidth}{!}{$
         W_h(q(x)) =  \frac{1}{\pi h^3} \left\{\begin{matrix}
@@
       \end{matrix}\right.
-        $}
     \end{equation}
@@
     \begin{equation} \label{eq:gradKernel}
-        \resizebox{0.88\linewidth}{!}{$
-        \partial W_h(q(x)) =  x \frac{1}{h \left \| x \right \|} \frac{1}{\pi h^3} \left\{\begin{matrix}
+        \partial W_h(q(x)) =  x \cdot \frac{1}{h \left \| x \right \|} \cdot \frac{1}{\pi h^3} \left\{\begin{matrix}
         - 3q + \frac{9}{4}q^2 & 0 \leqslant q < 1 \\ 
@@
 
-\subsection{Navier-Stokes}
-    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
-The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
+\subsection{Divergence-Free solver}
 
-    \begin{equation} \label{eq:nsDensity}
-        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
+\iffalse
+\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
+function correctDivergenceError
+    for all particles $i$ do
+        calculateDensityChange $\frac{Dp}{Dt}_i$
+    for all particles $i$ do //Scale $\color{codegreen}a_i$
+        $a_i = \frac{a_i}{\Delta t}$
+    while (t < $t_{max}$) do //Start simulation loop
+        adapt time step $\Delta{t}$ 
+        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
+            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
+        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
+        for all particles $i$ do //Update positions
+            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
+        for all particles $i$ do //Update neighbourhoods
+            find neighbourhoods $N_i(t + \Delta{t})$
+        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
+            compute densities $\rho_i(0)$
+            compute factors $\alpha_i(0)$
+        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
+\end{lstlisting}
+\fi
+
+The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in equation %\ref{eq:navier}.
+
+The solver will ensure that the divergence error is of an acceptable level defined by the user.
+The divergence error is suppressed by a combined symmetric pressure force $\vect{F_{j\leftarrow i}^p} $ over all particle $i$'s neighbours $j$, that is applied to the predicted velocity of particle $i$.
+The pressure force is determined by
+
+\begin{equation} \label{eq:pressure}
+\vect{F_i^p} = -\frac{m_i}{\rho_i}\nabla \vect{p_i},
 \end{equation}
 
-    \begin{equation} \label{eq:nsVelocity}
-        \frac{D\textbf{v}}{Dt} = \frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
+where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}.
+
+\begin{equation} \label{eq:dpdt}
+\frac{D\rho_i}{Dt} = \sum_j m_j (\vect{v_i} - \vect{v_j})\nabla \vect{W_{ij}}
 \end{equation}
 
-    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
-The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
-The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
-From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
-From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
-In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
-The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
-To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
+\begin{equation} \label{eq:kappa}
+\nabla \vect{p_i} = \kappa_i^v \nabla \rho_i = \kappa_i^v \sum_j m_j \nabla \vect{W_{ij}}
+\end{equation}
 
-\subsection{Divergence solver}
+Equation \ref{eq:dpdt} defines how the density change is computed and \ref{eq:kappa} computes the pressure gradient with the help of the stiffness parameter $\kappa_i^v$, where
+
+\begin{equation}
+\kappa_i^v = \frac{D\rho}{Dt} \cdot \frac{\rho_i}{\Delta t \cdot \alpha_i}
+\end{equation}
+
+and $\alpha_i$ is the scaling factor only dependant on the position explained in section \ref{section:alpha}.
+The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
+The velocity of particle $i$ is thus changed to:
+
+\begin{equation} \label{eq:pressuretotal}
+\vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} \cdot \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+\end{equation}
+
+After the symmetric pressure force has been applied to the new velocity should the density change in equation \ref{eq:dpdt} be equal or close to zero.
+If the error is too large will the process be iterated until desired result have been achieved.
+
 \subsection{Density solver}
 
+\subsection{Navier-stokes}
 % Write about non pressure forces, velocity osv..?
@@
 The quicker the particles move the smaller time steps are taken such that the numerical 
-error of the explicit time integration is negligible.
-For this is a condition proposed by Courant–Friedrichs–Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
-The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by the maximum velocity $\vect{V_{max}}$ in the scene.
+error of the explicit time integration is negligible. For this is a condition proposed by 
+Courant$-$Friedrichs$-$Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}. 
+The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
+the maximum velocity $\vect{V_{max}}$ in the scene.
 
@@
 
-\subsection{Density and alpha factors}
+\subsection{Density and alpha factors} \label{section:alpha}
 \subsection{Screen space fluid rendering}

### 2016-12-14 report/paper.tex

--- 2016-12-14 report/paper.tex
@@
 \subsection{Divergence solver} \label{divergenceSolver}
+
+\iffalse
+\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
+function correctDivergenceError
+    for all particles $i$ do
+        calculateDensityChange $\frac{Dp}{Dt}_i$
+    for all particles $i$ do //Scale $\color{codegreen}a_i$
+        $a_i = \frac{a_i}{\Delta t}$
+    while (t < $t_{max}$) do //Start simulation loop
+        adapt time step $\Delta{t}$ 
+        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
+            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
+        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
+        for all particles $i$ do //Update positions
+            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
+        for all particles $i$ do //Update neighbourhoods
+            find neighbourhoods $N_i(t + \Delta{t})$
+        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
+            compute densities $\rho_i(0)$
+            compute factors $\alpha_i(0)$
+        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
+\end{lstlisting}
+\fi
+
+The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in equation %\ref{eq:navier}.
+
+The solver will ensure that the divergence error is of an acceptable level defined by the user.
+The divergence error is suppressed by a combined symmetric pressure force $\vect{F_{j\leftarrow i}^p} $ over all particle $i$'s neighbours $j$, that is applied to the predicted velocity of particle $i$.
+The pressure force is determined by
+
+\begin{equation} \label{eq:pressure}
+\vect{F_i^p} = -\frac{m_i}{\rho_i}\nabla \vect{p_i},
+\end{equation}
+
+where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}.
+
+\begin{equation} \label{eq:dpdt}
+\frac{D\rho_i}{Dt} = \sum_j m_j (\vect{v_i} - \vect{v_j})\nabla \vect{W_{ij}}
+\end{equation}
+
+\begin{equation} \label{eq:kappa}
+\nabla \vect{p_i} = \kappa_i^v \nabla \rho_i = \kappa_i^v \sum_j m_j \nabla \vect{W_{ij}}
+\end{equation}
+
+Equation \ref{eq:dpdt} defines how the density change is computed and \ref{eq:kappa} computes the pressure gradient with the help of the stiffness parameter $\kappa_i^v$, where
+
+\begin{equation}
+\kappa_i^v = \frac{D\rho}{Dt} \cdot \frac{\rho_i}{\Delta t \cdot \alpha_i}
+\end{equation}
+
+and $\alpha_i$ is the scaling factor only dependant on the position explained in section \ref{section:alpha}.
+The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
+The velocity of particle $i$ is thus changed to:
+
+\begin{equation} \label{eq:pressuretotal}
+\vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} \cdot \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+\end{equation}
+
+After the symmetric pressure force has been applied to the new velocity should the density change in equation \ref{eq:dpdt} be equal or close to zero.
+If the error is too large will the process be iterated until desired result have been achieved.
+
 \subsection{Density solver} \label{densitySolver}
@@
 
-\subsection{Density and alpha factors}
+\subsection{Density and alpha factors} \label{section:alpha}
     The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}.

### 2016-12-14 report/paper.tex
Restructured method section

@@
 
+\subsection{Navier-Stokes}
+    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
+The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
+
+    \begin{equation} \label{eq:nsDensity}
+        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
+    \end{equation}
+
+    \begin{equation} \label{eq:nsVelocity}
+        \frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
+    \end{equation}
+
+    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
+The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
+The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
+From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
+From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
+In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
+The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
+To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
+Bridson~\cite{bridson} uses splitting on Equation~\ref{eq:nsVelocity} which induce $\frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p$ which will be used further on in the simulation.
+Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.
+
+\subsection{Adapted time step}
+
+% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
+In order to assure a stable simulation it is necessary to adapt the time step depending on the motion. 
+The quicker the particles move the smaller time steps are taken such that the numerical 
+error of the explicit time integration is negligible. For this is a condition proposed by 
+Courant$-$Friedrichs$-$Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}. 
+The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
+the maximum velocity $\vect{V_{max}}$ in the scene.
+
+\begin{equation}\label{eq:cfl}
+\Delta t \leq \frac{0.4 \cdot d}{\vect{V_{max}}}
+\end{equation}
+
+Each rendered frame however have a constant time step in order to ensure no speed change in the playback of the rendered video. 
+This is decided from the user side as a desired fps parameter. 
+This means that the simulation loop can iterate multiple times within each frame adaptively depending on the maximum velocity.
+
+\subsection{Density and alpha factors} \label{section:alpha}
+    The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}.
+This implies that neighbouring particles closer to the current particle $x_i$ will have a greater affect on the density.
+
+    \begin{equation}\label{eq:density}
+        \rho_i = \sum_j \frac{m_j}{\rho_j} \rho_j W_{ij} = \sum_j m_j W_{ij}
+    \end{equation}
+
 \subsection{Neighbourhood search}
@@
 
-\subsection{Navier-Stokes}
-    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
-The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
+\subsection{Density solver} \label{densitySolver}
+    This solver aims to minimize the density error caused by the numerical integration in the divergence solver.
+The density error is determined by comparing the actual density to the rest density, $\rho - \rho_0$.
+There are several density solvers presented in previous works, but the solver used in this project was presented by Bender et.
+al.~\cite{bender} which uses the fact that $\alpha_i$ has been calculated before the solver which reduces the computations in the solver.
+The solver will be active while the average of the predicted densities for every particle differs from the rest density with an error greater than a constant \textit{maxError}.
+For every particle a predicted density $\rho_i^*$ is calculated through an Euler integration step by Equation~\ref{eq:predDensity}.
 
-    \begin{equation} \label{eq:nsDensity}
-        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
+    \begin{equation} \label{eq:predDensity}
+        \resizebox{0.88\linewidth}{!}{$
+        \begin{matrix}
+        \rho_i^* = \rho_i + \Delta t \frac{D\rho_i}{Dt} \\ 
+        \\
+        \Leftrightarrow \\ 
+        \\
+        \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
+        (\textbf{v}_i^* - \textbf{v}_j^*) \bigtriangledown W_{ij}
+        \end{matrix}
+        $}
     \end{equation}
 
-    \begin{equation} \label{eq:nsVelocity}
-        \frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
+    Since the stiffness parameter $\kappa_i$ is derived from the pressure forces, updating the stiffness parameter will correct the density error.
+The stiffness parameter is determined by Equation~\ref{stiffnessDensity}.
+
+    \begin{equation} \label{stiffnessDensity}
+    \kappa_i = \frac{1}{\Delta t^2}(\rho_i^*-\rho_0)\alpha_i
     \end{equation}
 
-    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
-The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
-The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
-From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
-From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
-In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
-The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
-To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
-Bridson~\cite{bridson} uses splitting on Equation~\ref{eq:nsVelocity} which induce $\frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p$ which will be used further on in the simulation.
-Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.
+    The velocity is updated according to the the divergence solver. %Det här behöver kanske utvecklas, väntar med det tills stycket om divergence solvern är klar, innehåller samma ekvationer delvis
+
+% Write about non pressure forces, velocity osv..?
 
@@
 
-The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in equation %\ref{eq:navier}.
+The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in equation \ref{eq:nsDensity}.
 
-The solver will ensure that the divergence error is of an acceptable level defined by the user.
+The solver will ensure that the divergence error is smaller than an acceptable level defined by the user.
 The divergence error is suppressed by a combined symmetric pressure force $\vect{F_{j\leftarrow i}^p} $ over all particle $i$'s neighbours $j$, that is applied to the predicted velocity of particle $i$.
@@
 
-\subsection{Density solver} \label{densitySolver}
-    This solver aims to minimize the density error caused by the numerical integration in the divergence solver.
-The density error is determined by comparing the actual density to the rest density, $\rho - \rho_0$.
-There are several density solvers presented in previous works, but the solver used in this project was presented by Bender et.
-al.~\cite{bender} which uses the fact that $\alpha_i$ has been calculated before the solver which reduces the computations in the solver.
-The solver will be active while the average of the predicted densities for every particle differs from the rest density with an error greater than a constant \textit{maxError}.
-For every particle a predicted density $\rho_i^*$ is calculated through an Euler integration step by Equation~\ref{eq:predDensity}.
-
-    \begin{equation} \label{eq:predDensity}
-        \resizebox{0.88\linewidth}{!}{$
-        \begin{matrix}
-        \rho_i^* = \rho_i + \Delta t \frac{D\rho_i}{Dt} \\ 
-        \\
-        \Leftrightarrow \\ 
-        \\
-        \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
-        (\textbf{v}_i^* - \textbf{v}_j^*) \bigtriangledown W_{ij}
-        \end{matrix}
-        $}
-    \end{equation}
-
-    Since the stiffness parameter $\kappa_i$ is derived from the pressure forces, updating the stiffness parameter will correct the density error.
-The stiffness parameter is determined by Equation~\ref{stiffnessDensity}.
-
-    \begin{equation} \label{stiffnessDensity}
-    \kappa_i = \frac{1}{\Delta t^2}(\rho_i^*-\rho_0)\alpha_i
-    \end{equation}
-
-    The velocity is updated according to the the divergence solver. %Det här behöver kanske utvecklas, väntar med det tills stycket om divergence solvern är klar, innehåller samma ekvationer delvis
-
-% Write about non pressure forces, velocity osv..?
-\subsection{Adapted time step}
-
-% CFL condition, explain why its necessary and how its assured that each rendered frame has the same time step.
-In order to assure a stable simulation it is necessary to adapt the time step depending on the motion. 
-The quicker the particles move the smaller time steps are taken such that the numerical 
-error of the explicit time integration is negligible. For this is a condition proposed by 
-Courant$-$Friedrichs$-$Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}. 
-The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
-the maximum velocity $\vect{V_{max}}$ in the scene.
-
-\begin{equation}\label{eq:cfl}
-\Delta t \leq \frac{0.4 \cdot d}{\vect{V_{max}}}
-\end{equation}
-
-Each rendered frame however have a constant time step in order to ensure no speed change in the playback of the rendered video. 
-This is decided from the user side as a desired fps parameter. 
-This means that the simulation loop can iterate multiple times within each frame adaptively depending on the maximum velocity.
-
-\subsection{Density and alpha factors} \label{section:alpha}
-    The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}.
-This implies that neighbouring particles closer to the current particle $x_i$ will have a greater affect on the density.
-
-    \begin{equation}\label{eq:density}
-        \rho_i = \sum_j \frac{m_j}{\rho_j} \rho_j W_{ij} = \sum_j m_j W_{ij}
-    \end{equation}
-
 \subsection{Screen space fluid rendering}

### 2016-12-14 report/paper.tex
Explained the simulation algorithm

@@
 
-\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
+Listing \ref{lst:simulation} drafts the produced simulation after following Bender et al.~\cite{bender} proposed method in detail.
+Upon starting the application will the particle neighbourhood $N_i$ be determined with a cell list solution explained in section \ref{section:neighbour}.
+For each particle will the alpha $\alpha_i$ and density $\rho_i$ parameters computed, both of which are commonly reused variables and thus worth storing to reduce computation cost.
+This is done before entering the simulation loop as initial values are required.
+During the loop will all these parameters be updated as the simulation progresses.
+
+Once in the simulation at line $7$ will first the time step be determined through the CFL condition explained in section \ref{section:timestep}.
+The new velocities for each particle is then predicted through the included external forces which are propagated with ordinary euler integration.
+
+\begin{lstlisting}[caption={Simulation algorithm},label={lst:simulation}, language=C++, mathescape=true, escapeinside={(*}{*)}] 
 function performFluidSimulation
@@
 \end{lstlisting}
+
+The density solver uses the precomputed parameter $\alpha_i$ together with this prediction to compute the pressure forces in each neighbourhood such that it can correct the density error $\rho_i^* - \rho_i = 0$.
+This is further explained in \ref{densitySolver}.
+
+After the particle position $\vect{x_i}$ has been updated are the neighbourhoods and the $\alpha_i$ and $\rho_i$ factors recalculated to reflect the new positions.
+Lastly, in line $19$ the divergence solver, much like the density solver, calculates the pressure forces to eliminate its error.
+In this case the divergence error $\frac{D\rho_i}{Dt}$ is corrected to be equal to $0$.
 
@@
 
-\subsection{Adapted time step}
+\subsection{Adapted time step} \label{section:timestep}
 
@@
 
-\subsection{Density and alpha factors} \label{section:alpha}
-    The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}.
-This implies that neighbouring particles closer to the current particle $x_i$ will have a greater affect on the density.
-
-    \begin{equation}\label{eq:density}
-        \rho_i = \sum_j \frac{m_j}{\rho_j} \rho_j W_{ij} = \sum_j m_j W_{ij}
-    \end{equation}
-
-\subsection{Neighbourhood search}
+\subsection{Neighbourhood search} \label{section:neighbour}
     Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles neighbours.
@@
     It is important to use the same kernel function for both $W_h$ and $\nabla W_h$ to get the prediction and the correction step to be compatible to each other.
+
+\subsection{Density and alpha factors} \label{section:alpha}
+    The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}.
+This implies that neighbouring particles closer to the current particle $x_i$ will have a greater affect on the density.
+
+    \begin{equation}\label{eq:density}
+        \rho_i = \sum_j \frac{m_j}{\rho_j} \rho_j W_{ij} = \sum_j m_j W_{ij}
+    \end{equation}
 

### 2016-12-14 report/paper.tex
filled in some background work

--- 2016-12-14 report/paper.tex
@@
 
-    % Bridson
-    ~\cite{bridson}.
+    % Write about divergence free 
+    In the field of computer graphics a variety of methods for simulating incompressible fluids exist.
+    This section will give a brief overview of related approaches.
 
-    % Write about Benders article
+    Bender et al.~\cite{bender} proposed a method for a stable implicit SPH solution.
+    The method uses a combination of two pressure solvers which enforce a low volume compression below 0.01$\%$ and a divergence-free velocity field.
+    It can be seen as enforcing incompressibility on both position and velocity level.
+    The low compression is important for realistic physical behavior and a divergence-free state increases the stability of the simulation which reduces the number of solver iterations.
+
+    Another incompressible SPH method was proposed by Hu et al.~\cite{hu} for multiphase flows where the time integration step to obtain velocities is divided into two half steps.
+    This is done to enforce both zero density variaton condition and the velocity divergence-free condition at each full time step.
+    During the first half step density fluctuations are eliminated by altering the intermediate particle position.
+    During the second half step errors in velocity divergence are resolved by altering the intermediate particle velocity.
+
+    Ihmsen et al.~\cite{ihmsen2014implicit} proposed an implicit incompressible SPH method with a pressure poisson equation made by the combination of a symmetric SPH pressure force and a SPH discretization of the continuity equation.
+    The pressure poisson equation is used to compute the pressure so that the pressure forces correct the intermediate velocities to a divergence-free state.
+
+    A predictive corrective incompressible SPH method was proposed by Solenthaler et al.~\cite{solenthaler} based on the Lagrangian SPH model.
+    Incompressibility was enforced using a prediction correction scheme to determine the particle pressures.
+    To achieve this the information about the density fluctuation was actively propagated through the fluid and the pressure values where updated until the targeted density was satisfied.
+
+    %A ghost fluid approach for free surface and solid boundary conditions for a SPH fluid simulation what proposed by Schechter et al.~\cite{ghost}.
 
@@
 % Should be mentioned early since it is used in mostly of the algorithms
-\subsection{Kernel} \label{kernel}
+\subsection{Kernel}
     A kernel function is used to simulate how particle-particle interactions decrease with the distance between the current particle and its neighbours.
@@
 
-\subsection{Navier-Stokes}
-    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
-The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
+\subsection{Divergence solver}
+\subsection{Density solver}
 
-    \begin{equation} \label{eq:nsDensity}
-        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
-    \end{equation}
-
-    \begin{equation} \label{eq:nsVelocity}
-        \frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
-    \end{equation}
-
-    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively.
-The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
-The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
-From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
-From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
-In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
-The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
-To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
-Bridson~\cite{bridson} uses splitting on Equation~\ref{eq:nsVelocity} which induce $\frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p$ which will be used further on in the simulation.
-Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.
-
-\subsection{Divergence solver} \label{divergenceSolver}
-
-\iffalse
-\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
-function correctDivergenceError
-    for all particles $i$ do
-        calculateDensityChange $\frac{Dp}{Dt}_i$
-    for all particles $i$ do //Scale $\color{codegreen}a_i$
-        $a_i = \frac{a_i}{\Delta t}$
-    while (t < $t_{max}$) do //Start simulation loop
-        adapt time step $\Delta{t}$ 
-        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
-            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
-        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
-        for all particles $i$ do //Update positions
-            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
-        for all particles $i$ do //Update neighbourhoods
-            find neighbourhoods $N_i(t + \Delta{t})$
-        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
-            compute densities $\rho_i(0)$
-            compute factors $\alpha_i(0)$
-        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
-\end{lstlisting}
-\fi
-
-The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in equation %\ref{eq:navier}.
-
-The solver will ensure that the divergence error is of an acceptable level defined by the user.
-The divergence error is suppressed by a combined symmetric pressure force $\vect{F_{j\leftarrow i}^p} $ over all particle $i$'s neighbours $j$, that is applied to the predicted velocity of particle $i$.
-The pressure force is determined by
-
-\begin{equation} \label{eq:pressure}
-\vect{F_i^p} = -\frac{m_i}{\rho_i}\nabla \vect{p_i},
-\end{equation}
-
-where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}.
-
-\begin{equation} \label{eq:dpdt}
-\frac{D\rho_i}{Dt} = \sum_j m_j (\vect{v_i} - \vect{v_j})\nabla \vect{W_{ij}}
-\end{equation}
-
-\begin{equation} \label{eq:kappa}
-\nabla \vect{p_i} = \kappa_i^v \nabla \rho_i = \kappa_i^v \sum_j m_j \nabla \vect{W_{ij}}
-\end{equation}
-
-Equation \ref{eq:dpdt} defines how the density change is computed and \ref{eq:kappa} computes the pressure gradient with the help of the stiffness parameter $\kappa_i^v$, where
-
-\begin{equation}
-\kappa_i^v = \frac{D\rho}{Dt} \cdot \frac{\rho_i}{\Delta t \cdot \alpha_i}
-\end{equation}
-
-and $\alpha_i$ is the scaling factor only dependant on the position explained in section \ref{section:alpha}.
-The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
-The velocity of particle $i$ is thus changed to:
-
-\begin{equation} \label{eq:pressuretotal}
-\vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} \cdot \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
-\end{equation}
-
-After the symmetric pressure force has been applied to the new velocity should the density change in equation \ref{eq:dpdt} be equal or close to zero.
-If the error is too large will the process be iterated until desired result have been achieved.
-
-\subsection{Density solver} \label{densitySolver}
-    This solver aims to minimize the density error caused by the numerical integration in the divergence solver.
-The density error is determined by comparing the actual density to the rest density, $\rho - \rho_0$.
-There are several density solvers presented in previous works, but the solver used in this project was presented by Bender et.
-al.~\cite{bender} which uses the fact that $\alpha_i$ has been calculated before the solver which reduces the computations in the solver.
-The solver will be active while the average of the predicted densities for every particle differs from the rest density with an error greater than a constant \textit{maxError}.
-For every particle a predicted density $\rho_i^*$ is calculated through an Euler integration step by Equation~\ref{eq:predDensity}.
-
-    \begin{equation} \label{eq:predDensity}
-        \resizebox{0.88\linewidth}{!}{$
-        \begin{matrix}
-        \rho_i^* = \rho_i + \Delta t \frac{D\rho_i}{Dt} \\ 
-        \\
-        \Leftrightarrow \\ 
-        \\
-        \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
-        (\textbf{v}_i^* - \textbf{v}_j^*) \bigtriangledown W_{ij}
-        \end{matrix}
-        $}
-    \end{equation}
-
-    Since the stiffness parameter $\kappa_i$ is derived from the pressure forces, updating the stiffness parameter will correct the density error.
-The stiffness parameter is determined by Equation~\ref{stiffnessDensity}.
-
-    \begin{equation} \label{stiffnessDensity}
-    \kappa_i = \frac{1}{\Delta t^2}(\rho_i^*-\rho_0)\alpha_i
-    \end{equation}
-
-    The velocity is updated according to the the divergence solver. %Det här behöver kanske utvecklas, väntar med det tills stycket om divergence solvern är klar, innehåller samma ekvationer delvis
-
+\subsection{Navier-stokes}
 % Write about non pressure forces, velocity osv..?
@@
 The quicker the particles move the smaller time steps are taken such that the numerical 
-error of the explicit time integration is negligible. For this is a condition proposed by 
-Courant$-$Friedrichs$-$Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}. 
-The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
-the maximum velocity $\vect{V_{max}}$ in the scene.
+error of the explicit time integration is negligible.
+For this is a condition proposed by Courant–Friedrichs–Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
+The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by the maximum velocity $\vect{V_{max}}$ in the scene.
 
@@
 
-\subsection{Density and alpha factors} \label{section:alpha}
-    The density in a region of the fluid with a particle $x_i$ in the center is calculated through Equation~\ref{eq:density} where $\rho_i$ is the density with the current particle $x_i$ in center, $\rho_j$ is the density with a neighbouring particle $x_j$ in center, $m_j$ is the mass for a neighbour particle and $W_{ij}$ is the kernel function described in Section~\ref{kernel}.
-This implies that neighbouring particles closer to the current particle $x_i$ will have a greater affect on the density.
-
-    \begin{equation}\label{eq:density}
-        \rho_i = \sum_j \frac{m_j}{\rho_j} \rho_j W_{ij} = \sum_j m_j W_{ij}
-    \end{equation}
-
+\subsection{Density and alpha factors}
 \subsection{Screen space fluid rendering}

### 2016-12-14 report/paper.tex
since the order of density and divergence solver were changed a lot of equations had to be moved

--- 2016-12-14 report/paper.tex
@@
         \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
-        (\textbf{v}_i^* - \textbf{v}_j^*) \bigtriangledown W_{ij}
+        (\textbf{v}_i^* - \textbf{v}_j^*) \nabla W_{ij}
         \end{matrix}
@@
 
-    Since the stiffness parameter $\kappa_i$ is derived from the pressure forces, updating the stiffness parameter will correct the density error.
-The stiffness parameter is determined by Equation~\ref{stiffnessDensity}.
+    The pressure force of a particle is determined by
+
+    \begin{equation} \label{eq:pressure}
+        \vect{F_i^p} = -\frac{m_i}{\rho_i}\nabla \vect{p_i}
+    \end{equation}
+
+    where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}, Equation~\ref{eq:pressureGrad}.
+
+    \begin{equation} \label{eq:pressureGrad}
+        \nabla \vect{p_i} = \kappa_i^v \nabla \rho_i = \kappa_i^v \sum_j m_j \nabla \vect{W_{ij}}
+    \end{equation}
+
+    Furter on, the pressure forces $\vect{F_{j\leftarrow i}^p} $ that act from particle $i$ on all its neighbours $j$ is calculated through
+
+    \begin{equation} \label{eq:pressure2}
+        \vect{F_{j\leftarrow i}^p} = -\frac{m_i}{\rho_i} \kappa_i^v m_j \nabla W_{ij}
+    \end{equation}
+
+    where the stiffness parameter $\kappa_i$ is determined by Equation~\ref{stiffnessDensity}.
 
@@
 
-    The velocity is updated according to the the divergence solver. %Det här behöver kanske utvecklas, väntar med det tills stycket om divergence solvern är klar, innehåller samma ekvationer delvis
+    To conserve momentum it is required that all inner pressure forces sum up to zero such that $\vect{F_i^p} + \vect{F_{j\leftarrow i}^p} = 0$.
+A proposed by Bender et.
+al~\cite{bender} the solver computes the total force $\vect{F_{i,total}^p}$ as
+
+    \begin{equation} \label{eq:pressureTotal}
+        \resizebox{0.88\linewidth}{!}{$
+        \begin{matrix}
+            \vect{F_{i,total}^p} = \vect{F_i^p} + \sum_j \vect{F_{i\leftarrow j}^p} \\
+            \\
+            \Leftrightarrow \\ 
+            \\
+            \vect{F_{i,total}^p} = -m_i \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+        \end{matrix}
+        $}
+    \end{equation}
+
+    Newton's second law $F = ma = m\frac{dv}{dt}$ implies that
+
+    \begin{equation} \label{dvdt}
+        \frac{dv}{dt} = \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+    \end{equation}
+
+    The velocity of particle $i$ is thus changed to:
+
+    \begin{equation} \label{eq:velocity}
+        \vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+    \end{equation}
 
@@
 
-The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in equation \ref{eq:nsDensity}.
-
-The solver will ensure that the divergence error is smaller than an acceptable level defined by the user.
-The divergence error is suppressed by a combined symmetric pressure force $\vect{F_{j\leftarrow i}^p} $ over all particle $i$'s neighbours $j$, that is applied to the predicted velocity of particle $i$.
-The pressure force is determined by
-
-\begin{equation} \label{eq:pressure}
-\vect{F_i^p} = -\frac{m_i}{\rho_i}\nabla \vect{p_i},
-\end{equation}
-
-where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}.
+    The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in Equation \ref{eq:nsDensity}.
+\\ \\ Equation \ref{eq:dpdt} defines how the density change is computed.
+As in the density solver, the total pressure forces acting on a particle $i$ are computed.
+The stiffness parameter is calculated through~\ref{eq:kappaDivergence}.
 
@@
 
-\begin{equation} \label{eq:kappa}
-\nabla \vect{p_i} = \kappa_i^v \nabla \rho_i = \kappa_i^v \sum_j m_j \nabla \vect{W_{ij}}
-\end{equation}
-
-Equation \ref{eq:dpdt} defines how the density change is computed and \ref{eq:kappa} computes the pressure gradient with the help of the stiffness parameter $\kappa_i^v$, where
-
-\begin{equation}
+    \begin{equation}\label{eq:kappaDivergence}
         \kappa_i^v = \frac{D\rho}{Dt} \cdot \frac{\rho_i}{\Delta t \cdot \alpha_i}
@@
 
-and $\alpha_i$ is the scaling factor only dependant on the position explained in section \ref{section:alpha}.
     The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
-The velocity of particle $i$ is thus changed to:
-
-\begin{equation} \label{eq:pressuretotal}
-\vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} \cdot \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
-\end{equation}
-
-After the symmetric pressure force has been applied to the new velocity should the density change in equation \ref{eq:dpdt} be equal or close to zero.
-If the error is too large will the process be iterated until desired result have been achieved.
+After the symmetric pressure force has been applied to the new velocity should the density change in Equation \ref{eq:dpdt} be equal or close to zero.
+If the error is too large the process will be iterated until desired result have been achieved.
 

### 2016-12-14 report/paper.tex

@@
         \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
-        (\textbf{v}_i^* - \textbf{v}_j^*) \nabla W_{ij}
+        (\textbf{v}_i^* - \textbf{v}_j^*) \bigtriangledown W_{ij}
         \end{matrix}
@@
 
-    The pressure force of a particle is determined by
-
-    \begin{equation} \label{eq:pressure}
-        \vect{F_i^p} = -\frac{m_i}{\rho_i}\nabla \vect{p_i}
-    \end{equation}
-
-    where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}, Equation~\ref{eq:pressureGrad}.
-
-    \begin{equation} \label{eq:pressureGrad}
-        \nabla \vect{p_i} = \kappa_i^v \nabla \rho_i = \kappa_i^v \sum_j m_j \nabla \vect{W_{ij}}
-    \end{equation}
-
-    Furter on, the pressure forces $\vect{F_{j\leftarrow i}^p} $ that act from particle $i$ on all its neighbours $j$ is calculated through
-
-    \begin{equation} \label{eq:pressure2}
-        \vect{F_{j\leftarrow i}^p} = -\frac{m_i}{\rho_i} \kappa_i^v m_j \nabla W_{ij}
-    \end{equation}
-
-    where the stiffness parameter $\kappa_i$ is determined by Equation~\ref{stiffnessDensity}.
+    Since the stiffness parameter $\kappa_i$ is derived from the pressure forces, updating the stiffness parameter will correct the density error.
+The stiffness parameter is determined by Equation~\ref{stiffnessDensity}.
 
@@
 
-    To conserve momentum it is required that all inner pressure forces sum up to zero such that $\vect{F_i^p} + \vect{F_{j\leftarrow i}^p} = 0$.
-A proposed by Bender et.
-al~\cite{bender} the solver computes the total force $\vect{F_{i,total}^p}$ as
-
-    \begin{equation} \label{eq:pressureTotal}
-        \resizebox{0.88\linewidth}{!}{$
-        \begin{matrix}
-            \vect{F_{i,total}^p} = \vect{F_i^p} + \sum_j \vect{F_{i\leftarrow j}^p} \\
-            \\
-            \Leftrightarrow \\ 
-            \\
-            \vect{F_{i,total}^p} = -m_i \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
-        \end{matrix}
-        $}
-    \end{equation}
-
-    Newton's second law $F = ma = m\frac{dv}{dt}$ implies that
-
-    \begin{equation} \label{dvdt}
-        \frac{dv}{dt} = \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
-    \end{equation}
-
-    The velocity of particle $i$ is thus changed to:
-
-    \begin{equation} \label{eq:velocity}
-        \vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
-    \end{equation}
+    The velocity is updated according to the the divergence solver. %Det här behöver kanske utvecklas, väntar med det tills stycket om divergence solvern är klar, innehåller samma ekvationer delvis
 
@@
 
-    The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in Equation \ref{eq:nsDensity}.
-\\ \\ Equation \ref{eq:dpdt} defines how the density change is computed.
-As in the density solver, the total pressure forces acting on a particle $i$ are computed.
-The stiffness parameter is calculated through~\ref{eq:kappaDivergence}.
+The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in equation \ref{eq:nsDensity}.
+
+The solver will ensure that the divergence error is smaller than an acceptable level defined by the user.
+The divergence error is suppressed by a combined symmetric pressure force $\vect{F_{j\leftarrow i}^p} $ over all particle $i$'s neighbours $j$, that is applied to the predicted velocity of particle $i$.
+The pressure force is determined by
+
+\begin{equation} \label{eq:pressure}
+\vect{F_i^p} = -\frac{m_i}{\rho_i}\nabla \vect{p_i},
+\end{equation}
+
+where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}.
 
@@
 
-    \begin{equation}\label{eq:kappaDivergence}
+\begin{equation} \label{eq:kappa}
+\nabla \vect{p_i} = \kappa_i^v \nabla \rho_i = \kappa_i^v \sum_j m_j \nabla \vect{W_{ij}}
+\end{equation}
+
+Equation \ref{eq:dpdt} defines how the density change is computed and \ref{eq:kappa} computes the pressure gradient with the help of the stiffness parameter $\kappa_i^v$, where
+
+\begin{equation}
 \kappa_i^v = \frac{D\rho}{Dt} \cdot \frac{\rho_i}{\Delta t \cdot \alpha_i}
@@
 
+and $\alpha_i$ is the scaling factor only dependant on the position explained in section \ref{section:alpha}.
 The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
-After the symmetric pressure force has been applied to the new velocity should the density change in Equation \ref{eq:dpdt} be equal or close to zero.
-If the error is too large the process will be iterated until desired result have been achieved.
+The velocity of particle $i$ is thus changed to:
+
+\begin{equation} \label{eq:pressuretotal}
+\vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} \cdot \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+\end{equation}
+
+After the symmetric pressure force has been applied to the new velocity should the density change in equation \ref{eq:dpdt} be equal or close to zero.
+If the error is too large will the process be iterated until desired result have been achieved.
 

### 2016-12-14 report/paper.tex
hopefully finished with background

@@
 
-    %A ghost fluid approach for free surface and solid boundary conditions for a SPH fluid simulation what proposed by Schechter et al.~\cite{ghost}.
+    A ghost fluid approach for free surface and solid boundary conditions for a SPH fluid simulation what proposed by Schechter et al.~\cite{ghost}.
+    The ghost particles are placed on free surfaces and at solid boundary conditions.
+    The approach solves problems like spurious numerical surface tension artifacts and errors in the mass conservation constraints.
 

### 2016-12-15 report/paper.tex
alhpa factors

--- 2016-12-14 report/paper.tex
@@
     \begin{equation} \label{eq:nsDensity}
-        \frac{D\rho}{Dt} = 0 \Leftrightarrow \bigtriangledown \cdot \textbf{v} = 0 
+        \frac{D\rho}{Dt} = 0 \Leftrightarrow \nabla \cdot \textbf{v} = 0 
     \end{equation}
@@
     \begin{equation} \label{eq:nsVelocity}
-        \frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
+        \frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \nabla p + v \nabla ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
     \end{equation}
@@
 The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
-The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
-From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
+The second part of Equation~\ref{eq:nsDensity} $\nabla \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
+From the continuity equation $\frac{D\rho}{Dt} = -\rho \nabla \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
 From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
@@
 To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
-Bridson~\cite{bridson} uses splitting on Equation~\ref{eq:nsVelocity} which induce $\frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p$ which will be used further on in the simulation.
 Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.
@@
 
+    Both the density solver and the divergence solver uses a stiffness parameter $\kappa_i^v$ presented by Bender et. al.~\cite{bender}, Equation~\ref{eq:kappa}.
+
+    \begin{equation}\label{eq:kappa}
+        \kappa_i^v = \frac{1}{\Delta t} \frac{D\rho_i}{Dt} \cdot \alpha_i
+    \end{equation}
+
+    In the equation $\Delta t$ denotes a small time step and $\alpha_i$ is a precomputed factor to decrease the computations while performing the solvers.
+Bender et.
+al.~\cite{bender} defines $\alpha_i$ as
+
+    \begin{equation}\label{eq:alpha}
+        \alpha_i = \frac{\rho_i}{\left | \sum_j m_j \nabla \vect{W}_{ij}  \right |^2 + \sum_j \left | m_j \nabla \vect{W}_{ij} \right |^2}
+    \end{equation}
+
+    where $\vect{W}_{ij}$ corresponds to the kernel gradient.
+
 \subsection{Density solver} \label{densitySolver}
@@
         \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
-        (\textbf{v}_i^* - \textbf{v}_j^*) \nabla W_{ij}
+        (\textbf{v}_i^* - \textbf{v}_j^*) \nabla \vect{W}_{ij}
         \end{matrix}
@@
     \begin{equation} \label{eq:pressure2}
-        \vect{F_{j\leftarrow i}^p} = -\frac{m_i}{\rho_i} \kappa_i^v m_j \nabla W_{ij}
+        \vect{F_{j\leftarrow i}^p} = -\frac{m_i}{\rho_i} \kappa_i^v m_j \nabla \vect{W}_{ij}
     \end{equation}
@@
             \\
-            \vect{F_{i,total}^p} = -m_i \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+            \vect{F_{i,total}^p} = -m_i \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W}_{ij}
         \end{matrix}
@@
     \begin{equation} \label{dvdt}
-        \frac{dv}{dt} = \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+        \frac{dv}{dt} = \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W}_{ij}
     \end{equation}
@@
     \begin{equation} \label{eq:velocity}
-        \vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W_{ij}}
+        \vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W}_{ij}
     \end{equation}
@@
     \begin{equation} \label{eq:dpdt}
-        \frac{D\rho_i}{Dt} = \sum_j m_j (\vect{v_i} - \vect{v_j})\nabla \vect{W_{ij}}
+        \frac{D\rho_i}{Dt} = \sum_j m_j (\vect{v_i} - \vect{v_j})\nabla \vect{W}_{ij}
     \end{equation}
@@
     \begin{equation}\label{eq:kappaDivergence}
-        \kappa_i^v = \frac{D\rho}{Dt} \cdot \frac{\rho_i}{\Delta t \cdot \alpha_i}
+        \kappa_i^v = \frac{1}{\Delta t} \frac{D\rho_i}{Dt} \cdot \alpha_i
     \end{equation}

### 2016-12-15 report/paper.tex
started on implementation

--- 2016-12-14 report/paper.tex
@@
 %*No code*!!!!!
+    % Talk about what we used for plugins, computers etc maybe or should this be in method?
 
-    % Talk about what we used for plugins, computers etc maybe or should this be in method?
+    % Libraries
+    % Computers
+    % OpenGL
+    % language
+    % Makefiles
+    The divergence-free SPH simulation was developed using OpenGL and C++ together with the OpenGL mathematics library GLM for vector calculations.
+    FreeImage was used to save frames of the simulation and FFmpeg to assembled the frames into a video.
+    PicoJSON was used to load in variables from a file making it easier to change variables quick and efficient and a makefile was used to compile the project.
+
+    The simulation was developed on Mac OS and Linux.
+    The Linux computer is running on Debian and has four Intel(R) Core(TM) i5-2450M CPU @ 2.50GHz, one NVIDIA GeForce 610M graphics card and 6.0 GiB RAM.
 
@@
 
+    % Write about FPS when running x amount of particles
+
 \section{Conclusions and Future Work}
@@
 
+% Parallelize
+
 \bibliographystyle{ieeetr}

### 2016-12-15 report/paper.tex
dirichlet boundary conditions

--- 2016-12-15 report/paper.tex
@@
 
-\subsection{Screen space fluid rendering}
+\subsection{Boundary conditions}
+Dirichlet boundary condition was implemented in order to simulate the fluid inside a bounding box.
+Dirichlet assumes that the pressure is zero outside the fluid which and therefore the value a quantity has to take along the boundary has to be specified, in this case the velocity was set to zero when a particle reaches the boundary, Bridson~\cite{bridson}.
 

### 2016-12-15 report/paper.tex
abstract

@@
 \begin{onecolabstract}
-    \noindent This is the paper's abstract \ldots
-    \vspace{1em}\\{\bf Index Terms:} Divergence-free, SPH, divergence correction, density correction.\vspace{1em}
+    \noindent This paper discusses the technique presented by Bender et.
+al.~\cite{bender} for simulating incompressible fluids with an efficient and stable Divergence-Free Smoothed Particle Hydrodynamics method.
+The method uses two solvers to obtain a divergence-free fluid, a divergence solver and a density solver to correct the density after the divergence has been corrected.
+This results in a stable an fast simulation of a divergence-free fluid even though the simulation is performed on the CPU without any parallelization.
+    \vspace{1em}\\{\bf Index Terms:} Incompressible fluids, Divergence-free, SPH, divergence correction, density correction.\vspace{1em}
 \end{onecolabstract}
@@
     The Linux computer is running on Debian and has four Intel(R) Core(TM) i5-2450M CPU @ 2.50GHz, one NVIDIA GeForce 610M graphics card and 6.0 GiB RAM.
+One of the Mac OS computer is running on macOS Sierra and has a 1.8 GHz Intel Core i5 processor with an Intel HD Graphics 400 1536 MB graphics card.
 

### 2016-12-15 report/paper.tex
wrote about future work

--- 2016-12-15 report/paper.tex
@@
     % Use ghost-SPH or similar
+    To increase the accuracy of the interaction with the boundary and free surfaces a method like ghost-SPH that was presented by Schechter et al.~\cite{ghost} could be implemented.
+    A method that Bender et al.~\cite{bender} used was the versatile rigid fluid coupling method for incompressible SPH presented by Akinci et al.~\cite{akinci2012versatile}.
+    It is based on static boundary particles which could also be a valid implementation to get better boundary interactions.
 
     % Parallelize 
+    To get the simulation to be able to handle more particles the solution could be parallelized to speed up the calculations.
+    This could be done using OpenMP and increasing the threading.
+
+    % Viscosity
+    To get the simulation more interesting viscosity can be introduced into the fluid simulation.
+    It would make simulations of thicker fluids possible.
 

### 2016-12-15 report/paper.tex
boundary condition

--- 2016-12-15 report/paper.tex
@@
 Dirichlet boundary condition was implemented in order to simulate the fluid inside a bounding box.
-Dirichlet assumes that the pressure is zero outside the fluid which and therefore the value a quantity has to take along the boundary has to be specified, in this case the velocity was set to zero when a particle reaches the boundary, Bridson~\cite{bridson}.
+Dirichlet assumes that the pressure is zero outside the fluid and therefore the value a quantity has to take along the boundary has to be specified, in this case the velocity was set to zero when a particle reaches the boundary, Bridson~\cite{bridson}.
+
+An isosurface was also introduced in the middle of the scene in order to simulate collisions with arbitrary objects.
+This was done in an implicit representation of geometric shapes, as function $f(x,y,z)$.
+Given an isovalue $C = f(\vect{x})$, point $\vect{x}$ can be classified as follows:
+
+\begin{eqnarray}
+\label{eq:classifications}
+\begin{array}{ll}
+ \mbox{Inside:} & f(\mathbf{x}) < C \\
+ \mbox{Outside:} & f(\mathbf{x}) > C \\
+ \mbox{On surface:} & f(\mathbf{x}) = C \\  
+  \end{array}
+\end{eqnarray}
+
+To include all possible implicit geometries can a quadratic function be described as:
+
+\begin{eqnarray}
+\label{eqn:quadratic}
+f(x,y,z) &=& Ax^2+2Bxy+2Cxz\\
+\nonumber &+&2Dx+Ey^2+2Fyz\\
+\nonumber &+&2Gy+Hz^2+2Iz\\
+\nonumber &+&J
+\end{eqnarray}
+
+or on matrix form:
+
+\begin{eqnarray}
+\label{eqn:quadricmatrixform}
+\mathbf{p}^T\mathbf{Q}\mathbf{p} =  \left[ \begin{array}{cccc} x & y & z & 1 \end{array} \right] \left[ \begin{array}{cccc}
+A & B & C &D\\
+B & E & F & G\\
+C & F & H & I\\
+D & G & I & J\\
+ \end{array} \right] \left[ \begin{array}{c}x \\
+y \\
+z \\
+1 \end{array} \right]
+\end{eqnarray}
+
+Considering that the geometry is placed at the origin, $f(\vect{x}) = 0$ means the particle is on the surface and $\vect{p}^T\vect{Q}\vect{p} \leq 0$ can be used to determine collisions.
+
+Since the quadric surface is known analytically can the differentiation be applied to the quadric directly to get the gradient of the shape in position $\vect{x}$.
+This gives an efficient and analytical expression with a coefficient matrix $Q$, which is used to determine the new velocity direction of a particle after colliding with the surface.
+
+\begin{eqnarray}
+\nabla f(x,y,z)
+\label{eqn:quadricnormal}
+\nonumber = 2 \left[ \begin{array}{cccc}
+A & B & C &D\\
+B & E & F & G\\
+C & F & H & I\\
+ \end{array} \right] \left[ \begin{array}{c}x \\
+y \\
+z \\
+1 \end{array} \right]\\
+\nonumber = 2\mathbf{Q}_{sub}\mathbf{p}
+\end{eqnarray}
 
@@
 % Use ghost-SPH or similar
-    To increase the accuracy of the interaction with the boundary and free surfaces a method like ghost-SPH that was presented by Schechter et al.~\cite{ghost} could be implemented.
-    A method that Bender et al.~\cite{bender} used was the versatile rigid fluid coupling method for incompressible SPH presented by Akinci et al.~\cite{akinci2012versatile}.
-    It is based on static boundary particles which could also be a valid implementation to get better boundary interactions.
 
 % Parallelize
-    To get the simulation to be able to handle more particles the solution could be parallelized to speed up the calculations.
-    This could be done using OpenMP and increasing the threading.
-
-    % Viscosity
-    To get the simulation more interesting viscosity can be introduced into the fluid simulation.
-    It would make simulations of thicker fluids possible.
 

### 2016-12-15 report/paper.tex
fix to equation in boundary condition

@@
 \label{eqn:quadratic}
-f(x,y,z) &=& Ax^2+2Bxy+2Cxz\\
-\nonumber &+&2Dx+Ey^2+2Fyz\\
+f(x,y,z) &=
+\nonumber & Ax^2+2Bxy+2Cxz\\
+ &+&2Dx+Ey^2+2Fyz\\
 \nonumber &+&2Gy+Hz^2+2Iz\\
@@
 Since the quadric surface is known analytically can the differentiation be applied to the quadric directly to get the gradient of the shape in position $\vect{x}$.
-This gives an efficient and analytical expression with a coefficient matrix $Q$, which is used to determine the new velocity direction of a particle after colliding with the surface.
+This gives an efficient and analytical expression with a coefficient matrix $Q$, which is used to help determine the new velocity direction of a particle after colliding with the surface.
 
 \begin{eqnarray}
-\nabla f(x,y,z)
-\label{eqn:quadricnormal}
-\nonumber = 2 \left[ \begin{array}{cccc}
+\label{eq:quadricnormal}
+\nabla f(x,y,z) = 2 \left[ \begin{array}{cccc}
 A & B & C &D\\

### 2016-12-15 report/paper.tex
changed to one liner sentences and squashed some equations

@@
     Another incompressible SPH method was proposed by Hu et al.~\cite{hu} for multiphase flows where the time integration step to obtain velocities is divided into two half steps.
-    This is done to enforce both zero density variaton condition and the velocity divergence-free condition at each full time step.
+    This is done to enforce both zero density variation condition and the velocity divergence-free condition at each full time step.
     During the first half step density fluctuations are eliminated by altering the intermediate particle position.
@@
 The quicker the particles move the smaller time steps are taken such that the numerical 
-error of the explicit time integration is negligible. For this is a condition proposed by 
+error of the explicit time integration is negligible.
+For this is a condition proposed by 
 Courant$-$Friedrichs$-$Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
@@
 
-\begin{equation}\label{eq:cfl}
+\begin{equation} \label{eq:cfl}
 \Delta t \leq \frac{0.4 \cdot d}{\vect{V_{max}}}
@@
     In SPH simulations this is an approximation of the Gaussian kernel function.
-Different kernels have been tested in previous works i.e.
-the poly6 kernel, the spiky kernel and the cubic spline kernel.
-According to Bender et.
-al.~\cite{bender} the cubic spline kernel presented by Monaghan~\cite{monaghan} was used.
+    Different kernels have been tested in previous works i.e. the poly6 kernel, the spiky kernel and the cubic spline kernel.
+    According to Bender et. al.~\cite{bender} the cubic spline kernel presented by Monaghan~\cite{monaghan} was used.
     The kernel is described by Equation~\ref{eq:kernel}, where $q(x)=\frac{\left \| x \right \|}{h}$, $x$ is the distance between the current particle and a neighbour particle and $h$ is the support radius for the kernel.
@@
     The algorithm does also require the kernel gradient.
-To reduce the computational effort and memory requirements Bender et.
-al.~\cite{bender} introduce a scalar function $g(q) = \frac{\partial W_h}{\partial q} \cdot \frac{1}{h \left \| x \right \|}$.
+    To reduce the computational effort and memory requirements Bender et. al.~\cite{bender} introduce a scalar function $g(q) = \frac{\partial W_h}{\partial q} \cdot \frac{1}{h \left \| x \right \|}$.
     The gradient kernel is then calculated by $\partial W_h(q(x)) = x \cdot g(x)$.
@@
 
-    \begin{equation}\label{eq:density}
+    \begin{equation} \label{eq:density}
         \rho_i = \sum_j \frac{m_j}{\rho_j} \rho_j W_{ij} = \sum_j m_j W_{ij}
@@
 
-    \begin{equation}\label{eq:kappa}
+    \begin{equation} \label{eq:kappa}
         \kappa_i^v = \frac{1}{\Delta t} \frac{D\rho_i}{Dt} \cdot \alpha_i
@@
     In the equation $\Delta t$ denotes a small time step and $\alpha_i$ is a precomputed factor to decrease the computations while performing the solvers.
-Bender et.
-al.~\cite{bender} defines $\alpha_i$ as
+    Bender et. al.~\cite{bender} defines $\alpha_i$ as
 
-    \begin{equation}\label{eq:alpha}
+    \begin{equation} \label{eq:alpha}
         \alpha_i = \frac{\rho_i}{\left | \sum_j m_j \nabla \vect{W}_{ij}  \right |^2 + \sum_j \left | m_j \nabla \vect{W}_{ij} \right |^2}
@@
     \begin{equation} \label{eq:predDensity}
-        \resizebox{0.88\linewidth}{!}{$
+        \resizebox{0.85\linewidth}{!}{$
         \begin{matrix}
@@
 
-    Furter on, the pressure forces $\vect{F_{j\leftarrow i}^p} $ that act from particle $i$ on all its neighbours $j$ is calculated through
+    Further on, the pressure forces $\vect{F_{j\leftarrow i}^p} $ that act from particle $i$ on all its neighbours $j$ is calculated through
 
@@
     To conserve momentum it is required that all inner pressure forces sum up to zero such that $\vect{F_i^p} + \vect{F_{j\leftarrow i}^p} = 0$.
-A proposed by Bender et.
-al~\cite{bender} the solver computes the total force $\vect{F_{i,total}^p}$ as
+    A proposed by Bender et. al~\cite{bender} the solver computes the total force $\vect{F_{i,total}^p}$ as
 
     \begin{equation} \label{eq:pressureTotal}
-        \resizebox{0.88\linewidth}{!}{$
+        \resizebox{0.85\linewidth}{!}{$
         \begin{matrix}
@@
     \begin{equation} \label{eq:velocity}
+        \resizebox{0.85\linewidth}{!}{$
         \vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W}_{ij}
+        $}
     \end{equation}
@@
 
-\begin{eqnarray}
-\label{eq:classifications}
+\begin{eqnarray} \label{eq:classifications}
 \begin{array}{ll}
@@
 
-\begin{eqnarray}
-\label{eqn:quadricmatrixform}
+\begin{eqnarray} \label{eqn:quadricmatrixform}
+\resizebox{0.9\linewidth}{!}{$
 \mathbf{p}^T\mathbf{Q}\mathbf{p} =  \left[ \begin{array}{cccc} x & y & z & 1 \end{array} \right] \left[ \begin{array}{cccc}
@@
 1 \end{array} \right]
+$}
 \end{eqnarray}
@@
 
-\begin{eqnarray}
-\label{eq:quadricnormal}
+\begin{eqnarray} \label{eq:quadricnormal}
 \nabla f(x,y,z) = 2 \left[ \begin{array}{cccc}
@@
 
-    The simulation was developed on Mac OS and Linux.
+    The simulation was developed on macOS and Linux.
     The Linux computer is running on Debian and has four Intel(R) Core(TM) i5-2450M CPU @ 2.50GHz, one NVIDIA GeForce 610M graphics card and 6.0 GiB RAM.

### 2017-01-12 report/paper.tex
Errors corrected in report

@@
     The method uses two solvers to obtain a divergence-free fluid, a divergence solver and a density solver to correct the density after the divergence has been corrected.
-    This results in a stable an fast simulation of a divergence-free fluid even though the simulation is performed on the CPU without any parallelization.
+    This results in a stable and fast simulation of a divergence-free fluid even though the simulation is performed on the CPU without any parallelization.
     \vspace{1em}\\{\bf Index Terms:} Incompressible fluids, Divergence-free, SPH, divergence correction, density correction.\vspace{1em}
@@
 Listing \ref{lst:simulation} drafts the produced simulation after following Bender et al.~\cite{bender} proposed method in detail.
-Upon starting the application will the particle neighbourhood $N_i$ be determined with a cell list solution explained in section \ref{section:neighbour}.
-For each particle will the alpha $\alpha_i$ and density $\rho_i$ parameters computed, both of which are commonly reused variables and thus worth storing to reduce computation cost.
+Upon starting the application the particle neighbourhood $N_i$ will be determined with a cell list solution explained in section \ref{section:neighbour}.
+For each particle the alpha $\alpha_i$ and density $\rho_i$ parameters will be computed, both of which are commonly reused variables and thus worth storing to reduce computation cost.
 This is done before entering the simulation loop as initial values are required.
-During the loop will all these parameters be updated as the simulation progresses.
+During the loop all these parameters will be updated as the simulation progresses.
 
-Once in the simulation at line $7$ will first the time step be determined through the CFL condition explained in section \ref{section:timestep}.
+Once in the simulation at line $7$ the first time step will be determined through the CFL condition explained in section \ref{section:timestep}.
 The new velocities for each particle is then predicted through the included external forces which are propagated with ordinary euler integration.
@@
 
-After the particle position $\vect{x_i}$ has been updated are the neighbourhoods and the $\alpha_i$ and $\rho_i$ factors recalculated to reflect the new positions.
+After the particle position $\vect{x_i}$ has been updated the neighbourhoods and the $\alpha_i$ and $\rho_i$ factors are recalculated to reflect the new positions.
 Lastly, in line $19$ the divergence solver, much like the density solver, calculates the pressure forces to eliminate its error.
@@
     The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively.
-    The second part of Equation~\ref{eq:nsDensity} $\nabla \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence.
+    The second part of Equation~\ref{eq:nsDensity}, $\nabla \cdot \textbf{v} = 0$, induce that the velocity field is free from divergence.
     From the continuity equation $\frac{D\rho}{Dt} = -\rho \nabla \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}.
-    From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
+    From this it follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid.
     In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations.
@@
 This is decided from the user side as a desired fps parameter.
-This means that the simulation loop can iterate multiple times within each frame adaptively depending on the maximum velocity.
+This means that the simulation loop can iterate multiple times within each frame adaptively depending on the maximum velocity of a particle.
 
@@
 
-    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates and the fourth dimension is for storing the particles belonging to the cell.
+    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates and the fourth dimension is for storing the index of the particles that belong to the cell.
     The amount of cells are decided by dividing the scene into cells of length $h$.
-    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_min$ is the lowest position of the scene space.
+    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_{min}$ is the lowest position of the scene space.
     If the particle moves out of its cell it is reassigned to the new cell it is inside.
@@
 \subsection{Kernel} \label{kernel}
-    A kernel function is used to simulate how particle-particle interactions decrease with the distance between the current particle and its neighbours.
+    A kernel function is used to simulate how particle to particle interactions decrease with the distance between the current particle and its neighbours.
     In SPH simulations this is an approximation of the Gaussian kernel function.
@@
 
-    The pressure force of a particle is determined by
-
     \begin{equation} \label{eq:pressure}
@@
 
-    where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}, Equation~\ref{eq:pressureGrad}.
+    The pressure force of a particle is determined by Equation~\ref{eq:pressure}, where the pressure gradient can be solved using the SPH formulation proposed by Ihmsen et al.~\cite{ihmsen}, Equation~\ref{eq:pressureGrad}.
 
@@
 
-    Further on, the pressure forces $\vect{F_{j\leftarrow i}^p} $ that act from particle $i$ on all its neighbours $j$ is calculated through
-
     \begin{equation} \label{eq:pressure2}
@@
 
-    where the stiffness parameter $\kappa_i$ is determined by Equation~\ref{stiffnessDensity}.
+    Further on, the pressure forces $\vect{F_{j\leftarrow i}^p} $ that act from particle $i$ on all its neighbours $j$ is calculated through Equation~\ref{eq:pressure2}, where the stiffness parameter $\kappa_i$ is determined by Equation~\ref{stiffnessDensity}.
 
@@
     To conserve momentum it is required that all inner pressure forces sum up to zero such that $\vect{F_i^p} + \vect{F_{j\leftarrow i}^p} = 0$.
-    A proposed by Bender et. al~\cite{bender} the solver computes the total force $\vect{F_{i,total}^p}$ as
+    As proposed by Bender et. al~\cite{bender} the solver computes the total force $\vect{F_{i,total}^p}$ through Equation~\ref{eq:pressureTotal}.
 
@@
 
-    Newton's second law $F = ma = m\frac{dv}{dt}$ implies that
+    Newton's second law $F = ma = m\frac{dv}{dt}$ implies that $\frac{dv}{dt}$ is according to Equation~\ref{dvdt}. 
+    The velocity of particle $i$ is thus changed according to Equation~\ref{eq:velocity}.
 
@@
     \end{equation}
-
-    The velocity of particle $i$ is thus changed to:
 
@@
     As in the density solver, the total pressure forces acting on a particle $i$ are computed.
-    The stiffness parameter is calculated through~\ref{eq:kappaDivergence}.
+    The stiffness parameter is calculated through Equation~\ref{eq:kappaDivergence}.
 
@@
     The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
-    After the symmetric pressure force has been applied to the new velocity should the density change in Equation \ref{eq:dpdt} be equal or close to zero.
+    After the symmetric pressure force has been applied to the new velocity the density change in Equation \ref{eq:dpdt} should be equal or close to zero.
     If the error is too large the process will be iterated until desired result have been achieved.

### 2017-01-12 report/paper.tex
small changes in the report, added look-up-table to future work

--- 2016-12-15 report/paper.tex
@@
 \subsection{Navier-Stokes}
-    Navier-Stokes equations describes incompressible fluids and is therefore suitable for SPH simulations.
+    Navier-Stokes equations describe incompressible fluids and are therefore suitable for SPH simulations.
     The equations expressed in Lagrangian coordinates can be found in Equation~\ref{eq:nsDensity} and~\ref{eq:nsVelocity}.
@@
     To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
-    Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.
+    Section~\ref{densitySolver} and~\ref{divergenceSolver} give a deeper description of the solvers used in the project.
 
@@
 error of the explicit time integration is negligible.
-For this is a condition proposed by 
-Courant$-$Friedrichs$-$Lewy used often refered as the CFL-condition show in equation \ref{eq:cfl}.
+A condition was proposed by 
+Courant$-$Friedrichs$-$Lewy, often refered to as the CFL-condition, which is shown in Equation \ref{eq:cfl}.
 The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
-the maximum velocity $\vect{V_{max}}$ in the scene.
+the maximum velocity $\vect{V}_{max}$ in the scene.
 
 \begin{equation} \label{eq:cfl}
-\Delta t \leq \frac{0.4 \cdot d}{\vect{V_{max}}}
+\Delta t \leq \frac{0.4 \cdot d}{\vect{V}_{max}}
 \end{equation}
@@
 Each rendered frame however have a constant time step in order to ensure no speed change in the playback of the rendered video.
-This is decided from the user side as a desired fps parameter.
+This is decided from the user side as a desired fps, \textit{frames per second}, parameter.
 This means that the simulation loop can iterate multiple times within each frame adaptively depending on the maximum velocity.
@@
 \subsection{Neighbourhood search} \label{section:neighbour}
-    Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles neighbours.
+    Since SPH only considers a finite amount of neighbouring particles it is important to keep track of every particles' neighbours.
     Searching through all particles for neighbours within the cutoff distance $h$ for every particle is inefficient and takes $\mathcal{O}({N^2})$ time.
-    The cutoff distance $h$ is the kernels smoothing radius.
+    The cutoff distance $h$ is the kernels' smoothing radius.
     To make this faster a cell list was implemented.
@@
     A particle belongs to a cell if it is inside the space the cell occupy.
-    When finding the neighbour of particle $i$, only the neighbouring cells have to be searched for particles within the cutoff distance, see figure~\ref{fig:cellList}.
+    When finding the neighbour of particle $i$, only the neighbouring cells have to be searched for particles within the cutoff distance, see Figure~\ref{fig:cellList}.
 
@@
 
-    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates and the fourth dimension is for storing the particles belonging to the cell.
+    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the $x$, $y$ and $z$ coordinates and the fourth dimension is for storing the particles belonging to the cell.
     The amount of cells are decided by dividing the scene into cells of length $h$.
@@
         \\
-        \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
-        (\textbf{v}_i^* - \textbf{v}_j^*) \nabla \vect{W}_{ij}
+        \rho_i^* = \rho_i + \Delta t \sum_j m_j 
+        (\textbf{v}_{i}^* - \textbf{v}_j^*) \nabla \vect{W}_{ij}
         \end{matrix}
@@
         \resizebox{0.85\linewidth}{!}{$
-        \vect{v_i^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W}_{ij}
+        \vect{v_{i+1}^*}(t) = \vect{v_i^*}(t) - \Delta t \sum_j m_j (\frac{\kappa_i^v}{\rho_i} + \frac{\kappa_j^v}{\rho_i}) \nabla \vect{W}_{ij}
         $}
@@
 
-    The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
+    The stiffness parameters need to be determined iteratively since neighboring particles depend on each other.
     After the symmetric pressure force has been applied to the new velocity should the density change in Equation \ref{eq:dpdt} be equal or close to zero.
@@
 An isosurface was also introduced in the middle of the scene in order to simulate collisions with arbitrary objects.
-This was done in an implicit representation of geometric shapes, as function $f(x,y,z)$.
+This was done as an implicit representation of geometric shapes, described as a function $f(x,y,z)$.
 Given an isovalue $C = f(\vect{x})$, point $\vect{x}$ can be classified as follows:
@@
 
-To include all possible implicit geometries can a quadratic function be described as:
+All possible implicit geomitries can be descibed by a quadratic function, which is found in Equation~\ref{eqn:quadratic}, and in matrix form in Equation~\ref{eqn:quadricmatrixform}.
 
@@
 \end{eqnarray}
-
-or on matrix form:
 
@@
 
-Since the quadric surface is known analytically can the differentiation be applied to the quadric directly to get the gradient of the shape in position $\vect{x}$.
+Since the quadric surface is known analytically the differentiation can be applied to the quadric directly to get the gradient of the shape in position $\vect{x}$.
 This gives an efficient and analytical expression with a coefficient matrix $Q$, which is used to help determine the new velocity direction of a particle after colliding with the surface.
@@
 
+    % Look up table for kernel values
+    According to Bender et al.~\cite{bender} the kernel values could easily be precomputed and used together with a look up-table in order to speed up the calculations.
+This will result in a 30 $\%$ faster application.
+This could ether be done once when the iteration loop begins and the positions for the particles are determined.
+Another approach would be to compute a discrete representation of the kernel function before the iteration loop and interpolate between to kernel values in the simulation loop in order to find a kernel value for a specific particle.
+
 \bibliographystyle{ieeetr}

### 2017-01-12 particle/SPH.cpp
boundaryCondition update (ready for merge with master), need to implement cylinder vertex points correctly

@@ -31,83 +31,13 @@ SPH::SPH()
 	srand(time(0));
 	double step = 3.0 * particleRadius / boundaryDimension;
 
-	water = new vector<Particle*>(particleCount); //+(5*pow(ceil(1 + 2*boundaryDimension/step),2)));
+	water = new vector<Particle*>(particleCount);
     cellList = new CellList(
     	dvec3(-2*boundaryDimension, -2*boundaryDimension, -2*boundaryDimension), 
     	dvec3(2*boundaryDimension, 2*boundaryDimension+3.0, 2*boundaryDimension), H);
-    
-    // Create boundary particles
-    idx = 0;
-/*
-    // Bottom
-    randY = -boundaryDimension - particleRadius;
-    for (double i = -boundaryDimension-0.1; i < boundaryDimension+0.1; i+=step) {
-    	for (double j = -boundaryDimension-0.1; j < boundaryDimension+0.1; j+=step) {
-    		water->at(idx) = new Particle();
-			water->at(idx)->setPosition(i, randY, j);
-			water->at(idx)->setVelocity(dvec3(0.0, 0.0, 0.0));
-			water->at(idx)->id = 1;
-			
-			idx++;
-    	}
-    }
-
-    // Front
-    randZ = boundaryDimension + 0.1;
-    for (double i = -boundaryDimension-0.1; i < boundaryDimension+0.1; i+=step) {
-    	for (double j = -boundaryDimension-0.1; j < -boundaryDimension+0.2; j+=step) {
-    		water->at(idx) = new Particle();
-			water->at(idx)->setPosition(i, j, randZ);
-			water->at(idx)->setVelocity(dvec3(0.0, 0.0, 0.0));
-			water->at(idx)->id = 1;
-
-			idx++;
-    	}
-    }
-    // Back
-    randZ = -boundaryDimension - 0.1;
-    for (double i = -boundaryDimension-0.1; i < boundaryDimension+0.1; i+=step) {
-    	for (double j = -boundaryDimension-0.1; j < -boundaryDimension+0.2; j+=step) {
-    		water->at(idx) = new Particle();
-			water->at(idx)->setPosition(i, j, randZ);
-			water->at(idx)->setVelocity(dvec3(0.0, 0.0, 0.0));
-			water->at(idx)->id = 1;
-
-			idx++;
-    	}
-    }
-    // Left
-    randX = -boundaryDimension - 0.1;
[…]

### 2017-01-12 particle/SPH.h

@@ -36,7 +36,6 @@ class SPH
 
 	protected:
 		int				particleCount;
-		int 			idx;	
 		double			dT;
 		double 			restDensity;
 		double			avgDensityAdv;

### 2017-01-12 render/Cylinder.cpp

@@ -1,0 +1,110 @@
+/*************************************************************************
+Name:       Isabell Jansson, Jonathan Bosson, Ronja Grosz
+File name:  BoundingBox.cpp
+
+This class contains the data structure for creating and rendering
+a cylinder. 
+*************************************************************************/
+
+#include "Cylinder.h"
+
+Cylinder::Cylinder(){}
+
+Cylinder::~Cylinder() {}
+
+void Cylinder::draw(GLuint vao, float radius) {
+	GLfloat vertices[72] = {-radius, radius, radius, 	// 1
+							-radius, radius, -radius, // 2
+
+							-radius, radius, -radius, // 2
+							radius, radius, -radius, 	// 3
+
+							radius, radius, -radius, 	// 3
+							radius, radius, radius, 		// 4
+
+							radius, radius, radius, 		// 4
+							-radius, radius, radius, 	// 1
+
+							-radius, radius, radius, 	// 1
+							-radius, -radius, radius, 	// 5
+
+							-radius, -radius, radius, 	// 5
+							-radius, -radius, -radius, 	// 6
+
+							-radius, -radius, -radius, 	// 6
+							radius, -radius, -radius, 	// 7
+
+							radius, -radius, -radius, 	// 7
+							radius, -radius, radius, 	// 8
+
+							radius, -radius, radius, 	// 8
+							-radius, -radius, radius, 	// 5
+
+							-radius, radius, -radius, // 2
+							-radius, -radius, -radius, 	// 6
+
+							radius, radius, -radius, 	// 3
+							radius, -radius, -radius, 	// 7
+
+							radius, radius, radius, 		// 4
+							radius, -radius, radius}; 	// 8
+
+	GLfloat color[72] = {1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
+						1.0, 1.0, 1.0,
[…]

### 2017-01-12 render/Cylinder.h

@@ -1,0 +1,21 @@
+/*************************************************************************
+Name:	    Isabell Jansson, Jonathan Bosson, Ronja Grosz
+File name: 	Cylinder.h
+
+*************************************************************************/
+
+#include "GL/glew.h"
+#include "GLFW/glfw3.h"
+
+class Cylinder {
+	private:
+		int radius;
+		GLuint vbo_c[2];
+
+	public:
+		Cylinder();
+		~Cylinder();
+
+		void draw(GLuint vao, float radius);
+		
+};

### 2017-01-12 render/Viewport.cpp

@@ -243,7 +243,14 @@ int Viewport::start(int argc, char** argv) {
 	ifstream paramStream ("json/scene_parameters.json");
 	paramStream >> param;
 	float boundaryDimension = (float)(param.get<picojson::object>()["boundaryDimension"].get<double>());
+	float radius = (float)(param.get<picojson::object>()["geometry"].get<picojson::object>()["r"].get<double>());
 
+	/*
+		1 0 0 0 
+		0 1 0 0
+		0 0 0 0
+		0 0 0 r	
+	*/
 
     // Let's get started!
     while (!glfwWindowShouldClose(window)) {
@@ -278,7 +285,8 @@ int Viewport::start(int argc, char** argv) {
 
         
         boundingBox.draw(vao, boundaryDimension);
-        
+        cylinder.draw(vao, radius);
+
 		// Save the frame
 		if (record) {
 			frameCount++;

### 2017-01-12 render/Viewport.h

@@ -16,6 +16,7 @@ Viewport is used as a OpenGL controller.  Viewport is responsible for managing a
 
 #include "boost/timer.hpp"
 #include "../render/BoundingBox.h"
+#include "../render/Cylinder.h"
 
 #ifdef __linux__
 #include "../glm/glm/glm.hpp"
@@ -75,6 +76,7 @@ class Viewport {
 
 		SPH 		*hydro;
 		BoundingBox	boundingBox;
+		Cylinder	cylinder;
 		boost::timer		*timeSinceStart;
 
 		void init(void);					//Initializes a lot of openGL features, mostly just glEnable calls

### 2017-01-12 report/paper.tex

--- 2017-01-12 report/paper.tex
@@
 
-    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates and the fourth dimension is for storing the index of the particles that belong to the cell.
+    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the $x$, $y$ and $z$ coordinates and the fourth dimension is for storing the index of the particles that belong to the cell.
+
     The amount of cells are decided by dividing the scene into cells of length $h$.
@@
 
-    The stiffness parameters needs to be determined iteratively since neighboring particles depend on each other.
-    After the symmetric pressure force has been applied to the new velocity the density change in Equation \ref{eq:dpdt} should be equal or close to zero.
+    The stiffness parameters need to be determined iteratively since neighboring particles depend on each other.
+    After the symmetric pressure force has been applied to the new velocity should the density change in Equation \ref{eq:dpdt} be equal or close to zero.
     If the error is too large the process will be iterated until desired result have been achieved.

### 2017-01-12 report/paper.tex
Deleted a word

@@
     According to Bender et al.~\cite{bender} the kernel values could easily be precomputed and used together with a look up-table in order to speed up the calculations.
-This will result in a 30 $\%$ faster application.
-This could ether be done once when the iteration loop begins and the positions for the particles are determined.
+This will result in a $30\%$ faster application.
+This could be done once when the iteration loop begins and the positions for the particles are determined.
 Another approach would be to compute a discrete representation of the kernel function before the iteration loop and interpolate between to kernel values in the simulation loop in order to find a kernel value for a specific particle.

### 2017-01-12 report/paper.tex
results

@@
 
+When the application starts, the water is represented as set of spheres where every sphere has the properties of a particle.
+The initial state is illustrated Figure~\ref{fig:res1}.
+The water is affected by a gravity which induce a falling motion, illustrated in Figure~\ref{fig:res2}.
+
+\begin{figure}[h!]
+    \includegraphics[width=\linewidth]{img/frame0001.png}
+    \caption{Inital state of the water when the application starts.}
+    \label{fig:res1}
+\end{figure}
+
+\begin{figure}[h!]
+    \includegraphics[width=\linewidth]{img/frame0139.png}
+    \caption{The water falls due to the gravity.}
+    \label{fig:res2}
+\end{figure}
+
+Since the water aims to be in the state of rest, the motion will decrease over time which is illustrated in Figure~\ref{fig:res3}.
+\begin{figure}[h!]
+    \includegraphics[width=\linewidth]{img/frame0500.png}
+    \caption{caption}
+    \label{fig:res3}
+\end{figure}
+
+To be able to simulate how outer forces affect the water, we implemented a feature where the user is able to change the direction of the force.
+The result when applying a force from right to left on the water and then letting only the gravity affect it is shown in Figure~\ref{fig:res4}.
+\begin{figure}[h!]
+    \includegraphics[width=\linewidth]{img/frame0882.png}
+    \caption{caption}
+    \label{fig:res4}
+\end{figure}
+
+    % Results from adding cylinder
+
     % Write about FPS when running x amount of particles

### 2017-01-12 report/paper.tex
some corrections

--- 2017-01-12 report/paper.tex
@@
 
-    A ghost fluid approach for free surface and solid boundary conditions for a SPH fluid simulation what proposed by Schechter et al.~\cite{ghost}.
+    A ghost fluid approach for free surface and solid boundary conditions for a SPH fluid simulation was proposed by Schechter et al.~\cite{ghost}.
     The ghost particles are placed on free surfaces and at solid boundary conditions.
@@
 
-Listing \ref{lst:simulation} drafts the produced simulation after following Bender et al.~\cite{bender} proposed method in detail.
-Upon starting the application the particle neighbourhood $N_i$ will be determined with a cell list solution explained in section \ref{section:neighbour}.
+The produced simulation after following the method proposed by Bender et al.~\cite{bender} is drafted in Listing~\ref{lst:simulation}. 
+Upon starting the application the particle neighbourhood $N_i$ will be determined with a cell list solution explained in Section~\ref{section:neighbour}.
 For each particle the alpha $\alpha_i$ and density $\rho_i$ parameters will be computed, both of which are commonly reused variables and thus worth storing to reduce computation cost.
@@
 
-Once in the simulation at line $7$ the first time step will be determined through the CFL condition explained in section \ref{section:timestep}.
-The new velocities for each particle is then predicted through the included external forces which are propagated with ordinary euler integration.
+Once in the simulation at line $7$ the first time step will be determined through the CFL condition explained in Section~\ref{section:timestep}.
+The new velocities for each particle is then predicted through the included external forces which are propagated with ordinary Euler integration.
 
@@
 The density solver uses the precomputed parameter $\alpha_i$ together with this prediction to compute the pressure forces in each neighbourhood such that it can correct the density error $\rho_i^* - \rho_i = 0$.
-This is further explained in \ref{densitySolver}.
+This is further explained in Section~\ref{densitySolver}.
 
 After the particle position $\vect{x_i}$ has been updated the neighbourhoods and the $\alpha_i$ and $\rho_i$ factors are recalculated to reflect the new positions.
-Lastly, in line $19$ the divergence solver, much like the density solver, calculates the pressure forces to eliminate its error.
+Lastly, in line $19$ of Listing~\ref{lst:simulation} the divergence solver, much like the density solver, calculates the pressure forces to eliminate its error.
 In this case the divergence error $\frac{D\rho_i}{Dt}$ is corrected to be equal to $0$.
@@
 In order to assure a stable simulation it is necessary to adapt the time step depending on the motion.
-The quicker the particles move the smaller time steps are taken such that the numerical 
-error of the explicit time integration is negligible.
-A condition was proposed by 
-Courant$-$Friedrichs$-$Lewy, often refered to as the CFL-condition, which is shown in Equation \ref{eq:cfl}.
+Depending on how quick the particles move, smaller time steps are taken such that the numerical error of the explicit time integration is negligible.
+A condition was proposed by Courant$-$Friedrichs$-$Lewy, often refered to as the CFL-condition, which is shown in Equation~\ref{eq:cfl}.
 The time step $\Delta t$ is equal to a fraction of the particle's diameter $d$ divided by 
@@
     The amount of cells are decided by dividing the scene into cells of length $h$.
-    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation ~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_{min}$ is the lowest position of the scene space.
+    The particles are then assigned to a cell by finding the cell's coordinates that it belongs to according to Equation~\ref{eq:assignCell}, where $\vect{x}$ is the position of the particle and $\vect{s}_{min}$ is the lowest position of the scene space.
     If the particle moves out of its cell it is reassigned to the new cell it is inside.
@@
     The gradient kernel is then calculated by $\partial W_h(q(x)) = x \cdot g(x)$.
-    The gradient kernel is described by Equation ~\ref{eq:gradKernel}.
+    The gradient kernel is described by Equation~\ref{eq:gradKernel}.
 
@@
 
-    The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in Equation \ref{eq:nsDensity}.
-    \\ \\ Equation \ref{eq:dpdt} defines how the density change is computed.
+    The divergence-free solver ensures the fluid to be free of density change over time by enforcing $\frac{D\rho}{Dt} = 0$, which is equivalent to the Navier-Stokes divergence free condition given in Equation~\ref{eq:nsDensity}.
+    \\ \\ Equation~\ref{eq:dpdt} defines how the density change is computed.
     As in the density solver, the total pressure forces acting on a particle $i$ are computed.
@@
     The stiffness parameters need to be determined iteratively since neighboring particles depend on each other.
-    After the symmetric pressure force has been applied to the new velocity should the density change in Equation \ref{eq:dpdt} be equal or close to zero.
+    After the symmetric pressure force has been applied to the new velocity should the density change in Equation~\ref{eq:dpdt} be equal or close to zero.
     If the error is too large the process will be iterated until desired result have been achieved.

### 2017-01-12 report/paper.tex
Added some figure texts

@@
 	When the application starts, the water is represented as set of spheres where every sphere has the properties of a particle.
-The initial state is illustrated Figure~\ref{fig:res1}.
+	The initial state is illustrated in Figure~\ref{fig:res1}
 	The water is affected by a gravity which induce a falling motion, illustrated in Figure~\ref{fig:res2}.
@@
 		\includegraphics[width=\linewidth]{img/frame0500.png}
-    \caption{caption}
+		\caption{The water is calm with some motion.}
 		\label{fig:res3}
@@
 
-To be able to simulate how outer forces affect the water, we implemented a feature where the user is able to change the direction of the force.
+	To be able to simulate how outer forces affect the water, we implemented a feature where the user is able to change the direction of the force
 	The result when applying a force from right to left on the water and then letting only the gravity affect it is shown in Figure~\ref{fig:res4}.
+
 	\begin{figure}[h!]
 		\includegraphics[width=\linewidth]{img/frame0882.png}
-    \caption{caption}
+		\caption{The water after an applied right to left motion followed by gravity as the only outer force.}
 		\label{fig:res4}

### 2017-01-13 report/paper.tex
Started on sentence for fps
Added specs for frame rate
short conclusion

--- 2017-01-12 report/paper.tex
@@
 	The result when applying a force from right to left on the water and then letting only the gravity affect it is shown in Figure~\ref{fig:res4}.
+
 	\begin{figure}[h!]
 		\includegraphics[width=\linewidth]{img/frame0882.png}
-    \caption{caption}
+		\caption{The water after an applied right to left motion followed by gravity as the only outer force.}
 		\label{fig:res4}
@@
     % Write about FPS when running x amount of particles
+	When running the program in real time with 5000 particles it runs with a frame rate of 6-8 fps on the MAC computer.
+	When instead simulating with $10,000$ particles the frame rate is about 2 fps on the MAC computer.
 
@@
 %Suggest some ways that it could be extended and improved by adding more effort
+
+    % Conclusion
+    We were able to implement the SPH method presented by Bender et al.~\cite{bender} which induce that the fluid is incompressible.
+Though, there are several possible improvements that would improve the implementation and increase the frame rate.
 

### 2017-01-13 particle/SPH.cpp
rendered cylinder to show collision with implicit geometry works

@@ -40,9 +40,9 @@ SPH::SPH()
 	for(int i = 0; i < particleCount; ++i) {
 
 		// Random position
-		randX = ((float)rand()/(float)RAND_MAX) * 2.0 - 1.0;
-		randY = ((float)rand()/(float)RAND_MAX) * 2.0 - 0.0;
-		randZ = ((float)rand()/(float)RAND_MAX) * 2.0 - 1.0;
+		randX = (((float)rand()/(float)RAND_MAX) * 2.0 - 1.0) * boundaryDimension;
+		randY = (((float)rand()/(float)RAND_MAX) * 2.0 - 0.0) * boundaryDimension;
+		randZ = (((float)rand()/(float)RAND_MAX) * 2.0 - 1.0) * boundaryDimension;
 
 		if (alongBoundary(dvec4(randX, randY, randZ, 1.0)) != dvec3(0,0,0)) {
 			randY += 0.5;
@@ -193,8 +193,12 @@ void SPH::boundaryCondition(int i) {
 		vel.z = 0.0;
 
 	// Boundary of geometry
-	if (gradP != dvec3(0.0, 0.0, 0.0))
-		vel = gradP + gravity*dT;
+	if (gradP != dvec3(0.0, 0.0, 0.0)) {
+		if (vel.y > 0.0)
+			vel = dvec3(0.0, 0.0, 0.0);
+		else
+			vel = gradP + gravity*dT;
+	}
 
 	water->at(i)->setVelocity(vel);
 }

### 2017-01-13 render/Cylinder.cpp

@@ -7,72 +7,67 @@ a cylinder.
 *************************************************************************/
 
 #include "Cylinder.h"
+#include <math.h>
+#include <iostream>
+using namespace std;
 
 Cylinder::Cylinder(){}
 
-Cylinder::~Cylinder() {}
 
-void Cylinder::draw(GLuint vao, float radius) {
-	GLfloat vertices[72] = {-radius, radius, radius, 	// 1
-							-radius, radius, -radius, // 2
+Cylinder::~Cylinder(){}
 
-							-radius, radius, -radius, // 2
-							radius, radius, -radius, 	// 3
+void Cylinder::draw(GLuint vao, float radius, float boundary) {
 
-							radius, radius, -radius, 	// 3
-							radius, radius, radius, 		// 4
 
-							radius, radius, radius, 		// 4
-							-radius, radius, radius, 	// 1
+	// o=========o
 
-							-radius, radius, radius, 	// 1
-							-radius, -radius, radius, 	// 5
+	// how many segments? - 36
+	float segments = 200;
+	int l = 6;
+	const int num = l * segments;
+	float angle;
+	float x, y;
+	
+	GLfloat vertices[num];
+	float sqrtRadius = sqrt(radius);
+	for (int i = 0; i < segments; ++i) {
+		angle = i * 2/segments * M_PI;	
+		x = sqrtRadius * cos(angle);
+		y = sqrtRadius * sin(angle);
 
-							-radius, -radius, radius, 	// 5
-							-radius, -radius, -radius, 	// 6
+		// Lines on sides
+		vertices[l*i+0] = x;
+		vertices[l*i+1] = y;
+		vertices[l*i+2] = -boundary - 0.5;
 
-							-radius, -radius, -radius, 	// 6
-							radius, -radius, -radius, 	// 7
+		vertices[l*i+3] = x;
+		vertices[l*i+4] = y;
+		vertices[l*i+5] = boundary + 0.5;
 
-							radius, -radius, -radius, 	// 7
-							radius, -radius, radius, 	// 8
+		/*
+		// Back
+		vertices[l*i+6] = x;
+		vertices[l*i+7] = y;
+		vertices[l*i+8] = -boundary;
 
[…]

### 2017-01-13 render/Cylinder.h

@@ -8,14 +8,13 @@ File name: 	Cylinder.h
 #include "GLFW/glfw3.h"
 
 class Cylinder {
-	private:
-		int radius;
-		GLuint vbo_c[2];
+private:
+	int radius;
+	GLuint vbo_c[2];
 
-	public:
-		Cylinder();
-		~Cylinder();
+public:
+	Cylinder();
+	~Cylinder();
 
-		void draw(GLuint vao, float radius);
-		
+	void draw(GLuint vao, float radius, float boundary);
 };

### 2017-01-13 render/Viewport.cpp

@@ -244,6 +244,7 @@ int Viewport::start(int argc, char** argv) {
 	paramStream >> param;
 	float boundaryDimension = (float)(param.get<picojson::object>()["boundaryDimension"].get<double>());
 	float radius = (float)(param.get<picojson::object>()["geometry"].get<picojson::object>()["r"].get<double>());
+	string name = param.get<picojson::object>()["sceneName"].get<std::string>();
 
 	/*
 		1 0 0 0 
@@ -285,7 +286,12 @@ int Viewport::start(int argc, char** argv) {
 
         
         boundingBox.draw(vao, boundaryDimension);
-        cylinder.draw(vao, radius);
+        if (name == "cylinder")
+        	cylinder.draw(vao, radius, boundaryDimension);
+
+
+
+
 
 		// Save the frame
 		if (record) {

### 2017-01-13 render/Viewport.h

@@ -36,13 +36,6 @@ class SPH;
 
 using namespace std;
 
-struct rect {
-	float width;
-	float height;
-	float x;
-	float y;
-};
-
 class Viewport {	
 	public:	
 		Viewport();
@@ -76,8 +69,7 @@ class Viewport {
 
 		SPH 		*hydro;
 		BoundingBox	boundingBox;
-		Cylinder	cylinder;
-		boost::timer		*timeSinceStart;
+		Cylinder cylinder;
 
 		void init(void);					//Initializes a lot of openGL features, mostly just glEnable calls
 		void setupPerspective(GLFWwindow *window, GLfloat *P);			//OpenGL window reshape callback

### 2017-01-13 particle/SPH.cpp
better boundaryconditiot

@@ -194,7 +194,7 @@ void SPH::boundaryCondition(int i) {
 
 	// Boundary of geometry
 	if (gradP != dvec3(0.0, 0.0, 0.0)) {
-		if (vel.y > 0.0)
+		if (vel.y > 0.0 && pos.y < 0.0)
 			vel = dvec3(0.0, 0.0, 0.0);
 		else
 			vel = gradP + gravity*dT;

### 2017-01-15 particle/SPH.cpp
Include makefile for release verison, merge.sh now makes vid in realistic speed, cylinder collision is now on point

@@ -204,7 +204,7 @@ void SPH::boundaryCondition(int i) {
 }
 
 // Check if (x,y,z) is inside an implicit geometry, if so return the gradient in that p
-dvec3 SPH::alongBoundary(dvec4 p) {
+dvec3 SPH::alongBoundary(dvec4 p) { 
 	dmat4 Q = dmat4(0.0);
 	dvec4 gradP = dvec4(0.0, 0.0, 0.0, 0.0);
 
@@ -212,9 +212,10 @@ dvec3 SPH::alongBoundary(dvec4 p) {
 	if (sceneName == "cylinder") {
 		Q[0][0] = 1.0;
 		Q[1][1] = 1.0;
-		Q[3][3] = -geometry.w;
+		Q[3][3] = -pow((sqrt(geometry.w) + 2*particleRadius),2);
 	}
-
+	// g.w = r^2
+	// new g.w = r+particleRadius
 	// Ellipsoid
 	else if (sceneName == "ellipsoid") {
 		Q[0][0] = 1.0/pow(geometry.x,2);

### 2017-01-15 results/merge.sh

@@ -3,7 +3,7 @@
 # -i = image directory, %04d makes 0001 to 9999
 # -crf = quality, 15-25 is good
 
-ffmpeg -r 100 -f image2 -s 640x480 -i frames/frame%04d.png -vcodec libx264 -crf 25  -pix_fmt yuv420p out.mp4;
+ffmpeg -r 50 -f image2 -s 640x480 -i frames/frame%04d.png -vcodec libx264 -crf 25  -pix_fmt yuv420p out.mp4;
 
 cd frames/;
 rm *.png;

### 2017-01-15 report/paper.tex
added cylinder collision fig and text

@@
 	When the application starts, the water is represented as set of spheres where every sphere has the properties of a particle.
-	The initial state is illustrated in Figure~\ref{fig:res1}
+	The initial state is illustrated in Figure~\ref{fig:res1}.
 	The water is affected by a gravity which induce a falling motion, illustrated in Figure~\ref{fig:res2}.
@@
     % Results from adding cylinder
+    Implicit geometries can also be added to the scene in order to see how the water collides and interacts with it.
+Figure \ref{fig:res5} illustrates a scene with a solid cylinder that forces the water to divert its momentum around the cylinder's surface.
+
+    \begin{figure}[h!]
+        \includegraphics[width=\linewidth]{img/cylinder2.png}
+        \caption{The water colliding with a cylinder.}
+        \label{fig:res5}
+    \end{figure}
 

### 2017-01-17 report/paper.tex
last changes

@@
 
-    Ihmsen et al.~\cite{ihmsen2014implicit} proposed an implicit incompressible SPH method with a pressure poisson equation made by the combination of a symmetric SPH pressure force and a SPH discretization of the continuity equation.
-    The pressure poisson equation is used to compute the pressure so that the pressure forces correct the intermediate velocities to a divergence-free state.
+    Ihmsen et al.~\cite{ihmsen2014implicit} proposed an implicit incompressible SPH method with a pressure Poisson equation made by the combination of a symmetric SPH pressure force and a SPH discretization of the continuity equation.
+    The pressure Poisson equation is used to compute the pressure so that the pressure forces correct the intermediate velocities to a divergence-free state.
 
@@
     Incompressibility was enforced using a prediction correction scheme to determine the particle pressures.
-    To achieve this the information about the density fluctuation was actively propagated through the fluid and the pressure values where updated until the targeted density was satisfied.
+    To achieve this the information about the density fluctuation was actively propagated through the fluid and the pressure values were updated until the targeted density was satisfied.
 
@@
     The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions.
-    To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}.
+    To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fulfilled~\cite{bender}.
     Section~\ref{densitySolver} and~\ref{divergenceSolver} give a deeper description of the solvers used in the project.
@@
     After the symmetric pressure force has been applied to the new velocity should the density change in Equation~\ref{eq:dpdt} be equal or close to zero.
-    If the error is too large the process will be iterated until desired result have been achieved.
+    If the error is too large the process will be iterated until the desired result have been achieved.
 
@@
 
-All possible implicit geomitries can be descibed by a quadratic function, which is found in Equation~\ref{eqn:quadratic}, and in matrix form in Equation~\ref{eqn:quadricmatrixform}.
+All possible implicit geometries can be described by a quadratic function, which is found in Equation~\ref{eqn:quadratic}, and in matrix form in Equation~\ref{eqn:quadricmatrixform}.
 
@@
     FreeImage was used to save frames of the simulation and FFmpeg to assembled the frames into a video.
-    PicoJSON was used to load in variables from a file making it easier to change variables quick and efficient and a makefile was used to compile the project.
+    PicoJSON was used to load variables from a file making it easier to change variables quickly and efficiently and a makefile was used to compile the project.
 
     The simulation was developed on macOS and Linux.
-    The Linux computer is running on Debian and has four Intel(R) Core(TM) i5-2450M CPU @ 2.50GHz, one NVIDIA GeForce 610M graphics card and 6.0 GiB RAM.
+    The Linux computer is running on Debian and has a 50 GHz Intel Core i5-2450M processor and a NVIDIA GeForce 610M graphics card.
     One of the Mac OS computer is running on macOS Sierra and has a 1.8 GHz Intel Core i5 processor with an Intel HD Graphics 400 1536 MB graphics card.
@@
 		\includegraphics[width=\linewidth]{img/frame0001.png}
-		\caption{Inital state of the water when the application starts.}
+		\caption{Initial state of the water when the application starts.}
 		\label{fig:res1}
